### 2024-01-30 main.tex
First push

\documentclass{rapport}

\usepackage{myTools}

\usepackage{macros}

\author{Davide Fissore}
\title{Report - HO unification in Elpi} 
\date{January 29, 2024}

\xspaceaddexceptions{]\}}

\begin{document}

\maketitle

\def\elpi{\proglang{elpi}}
\def\coqelpi{\proglang{coq-elpi}}
\def\lambdaprolog{\proglang{$\lambda$-prolog}}
\def\coq{\proglang{coq}}
\newcommand*{\eqtau}{\ensuremath{\mathrel{\overset{\mathrm{\tau}}{=}}}}

\section{Higher-order unification in \texorpdfstring{$L_\lambda$}{L\_lambda}}

Higher-order unification is a unification problem known to be undecidable. The
goal is to determine a most general unifier between two lambda terms, where at
least one of contain a sub-term $F$ applied to a non empty list of arguments
$a_1,\dots,a_n$ and $F$ is a variable. Some researches have been done in order to
identify some easy patterns where it is possible to solve this kind of
unification.

In \cite{miller1991}, D. Miller provides a unification algorithm (the function
$\eqtau$) for higher-order variable when they are in the so-called pattern
fragment.

\section{From the object to the target language}

Higher order unification becomes difficult when we try to unify terms
represented in the syntax of an object language embedded inside another
language. This kind of interaction between different languages is something that
happens quite often in \coqelpi, a \coq plugin embedding \elpi.

\elpi is a variant of \lambdaprolog equipped with constraint handling rules. On
the other hand, \coq is a proof assistant used to reason about mathematical
properties and certify them thanks to its typing system.

Both languages allows to quantify higher-order variables and are able to perform
higher-order unification when needed. However, when we want to reason with the
\coq unification inside \elpi, it is less evident to reproduce the unification
between two terms. With this paper, we want to provide a framework explaining
how the unification of an object language (\coq in our case) can be expressed in
the targeted language (\elpi in our case).

Without any loss of generality, in \cref{table:coq2elpi} we present a subset of
the terms in the object language and for each term we give its representation in
the targeted language.

\begin{table}
  \centering
  \begin{tabular}{c||c|c}
  Term        & Obj. lang.                    & Targ. lang.                                   \\
  \hline\hline
  Constant    & a                             & `a'                                     \\
  Application & f a\_1 a\_2 ... a\_n          & app{[}`f', `a\_1', `a\_2', ..., `a\_n'{]}     \\
  Abstraction & fun (x : T) $\Rightarrow$ f x & fun `x' T (x\textbackslash app{[}`f', `x'{]}) \\
  Variable    & X                             & X                                             \\
  \end{tabular}
  \caption{Obj. lang. terms to targ. lang. terms representation}
  \label{table:coq2elpi}
\end{table}

\section{First-order unification}

Just as an introduction, we briefly show some small example of unification
between terms with only first-order unification variables. This way, we would
like the reader to become familiar between the communication of the two
languages.

Let's take as an example the following unification problem in the obj. lang.:

\begin{equation}
  f\ x\ 1 \eqtau f\ Y\ Z
  \label{eq:unif-fo-coq}
\end{equation}

\noindent where $f$, $x$ and $1$ are defined constant and $Y$ and $Z$ are both
unification variables. By convention we use upper case letter for quantified
variables. Moreover, for this first representation we do not really focus on the
type of the manipulated objects, since they do not condition the unification
algorithm.

It is quite evident that a valid substitution for \cref{eq:unif-fo-coq} is $\theta =
\{Y \mapsto `x', Z \mapsto `1'\}$. Now let's consider the same problem translated in
the targeted language.

\begin{equation}
  app[`f', `x', `1'] \eqtau app[`f', Y, Z]
  \label{eq:unif-fo-elpi}
\end{equation}

The unification of these terms is again quite simple since it is sufficient to
do a simple matching sub-term by sub-term so that variables can be instantiated.
We can therefore note that the same substitution $\theta$ will be produced.

\section{Higher-order unification}

The unification problem treated before was enough easy to be correctly
understood by both language representation. We want now to go a bit further and
reason with a more complex problem where a variable is a function of
higher-order.

We propose two different higher-order unification problem in the following
equations where, in the former we have rigid-flexible unification and in the
latter we have a flexible-flexible unification.

\begin{equation}
  f\ x\ 1 \eqtau F\ x
  \label{eq:unif-ho-coq1}
\end{equation}

\begin{equation}
  G\ x\ y \eqtau H\ y\ x
  \label{eq:unif-ho-coq2}
\end{equation}

The two substitutions for the previous examples are $\theta_1 = \{F \mapsto fun\
x \Rightarrow f\ x\ 1\}$ and $\theta_2 = \{H \mapsto fun\ y\ x \Rightarrow G\ x\
y\}$. We can note that to be in the pattern fragment, a functional variable
should be applied to distinct names.

If we translate the problem before in the target language, the unification
problems showed above become

\begin{equation}
  app[`f',\ `x',\ `1'] \eqtau app[F,\ `x']
  \label{eq:unif-ho-elpi1}
\end{equation}

\begin{equation}
  app[G,\ `x',\ `y'] \eqtau app[H,\ `y',\ `x']
  \label{eq:unif-ho-elpi2}
\end{equation}

Now, the new unification problems are no more expressed in the logic of the
targeted language and, therefore, in both cases, unification fails. The
procedure we can adopt in order to transform a higher-order unification problem
of the object language into the logic of the targeted language is to transform
the entry of the problem in a problem which can be understood by the targeted
language. The procedure is made of two steps:

\begin{enumerate}
  \item In the first place, we need to recognize the structure of the pattern
        fragment expressed in the term received in entry. This means that we
        need to find all the sub-terms of the form `app[X | L]', where `X' is
        a flexible variable and `L' is a list of distinct names.
  \item For any sub-term representing a higher-order unification in the object 
        language, we build a fresh variable `X\texttt{'}' such that the names 
        `L' are not in the scope of `X\texttt{'}', we call `X\texttt{'}' the 
        twin variable of `X'.
  \item We solve the new goal where each pattern fragment problem is replaced
        with a problem using twin variables and after each of these problems, we
        add a new premise linking these twin variables. The linking is done
        using the following criteria: for each abstraction in the resulting term
        `X\texttt{'}', unify recursively `X' to a lambda abstraction in the
        object language.
\end{enumerate}

The previous algorithm can be applied to \cref{eq:unif-ho-coq1,eq:unif-ho-coq2}
to provide the wanting solution. In particular, \cref{eq:unif-ho-coq1} is
transformed into the unification problem:

\begin{equation}
  f\ x\ 1 \eqtau F'\ x,\ ho-link\ F'\ F
  \label{eq:unif-ho-elpi1'}
\end{equation}

\begin{equation}
  G'\ x\ y \eqtau H'\ y\ x, ho-link\ G'\ G, ho-link\ H'\ H
  \label{eq:unif-ho-elpi2'}
\end{equation}

For instance, the former unification problem produce the substitution $\theta_1
= \{F' \mapsto (x\backslash\ f\ x\ 1)\}$. The ho-link function is then applied
to transform the substitution of $F'$ into the corresponding term of the object
langue: $F \mapsto fun \_\ \_\ (x\backslash\ f\ x\ 1)$ which correspond to the
term $fun\ x \Rightarrow f\ x\ 1$. The latter unification problem gives the
substitution $\theta_2 = \{H' \mapsto (y\ x \backslash G'\ x\ y)\}$ in the
targeted language. The first ho-link simply unify $G$ to $G'$ since $G'$ is
flexible, whereas $H$ is mapped to $fun\ y\ x \Rightarrow G'\ x\ y$.

The role of the ho-link is not only to instantiate the higher-order variable F
of the object language when F is flexible and the twin variable in the targeted
language is rigid. It may happen that F has already been partially instantiated.
The unification problem below gives such an example in the object language:

\begin{equation}
  G\ x\ y \eqtau H\ y\ x, H\ x\ y \eqtau x 
  \label{eq:unif-ho-coq4}
\end{equation}

\noindent producing the following substitution $\theta = \{G \mapsto (fun\ x\
y\Rightarrow x); F \mapsto (fun\ x\ y\Rightarrow y)\}$. This unification problem
is translated into:

\begin{equation}
  \begin{split}
    G'\ x\ y &\eqtau H'\ y\ x,\ ho-link\ G'\ G,\ ho-link\ H'\ H\\
    H''\ x\ y &\eqtau x,\ ho-link\ H''\ H
  \end{split}
  \label{eq:unif-ho-elpi4}
\end{equation}

\noindent The first line produces the same substitution as before: $\theta = \{H'
\mapsto (x\ y \backslash G'\ y\ x)\}$, where $G'$ is unified with $G$ and $H$ is
instantiated to <<$fun\ x\ y \Rightarrow G'\ y\ x$>>. While executing the second
line, we see $H''$ instantiated to the function $(x\ y\backslash x)$ in the
targeted language. The last ho-link is charged to link $H''$ with $H$ but this
time $H$ has already been partially instantiated. In particular the call to this
ho-link is as follows: $ho-link\ (x\ y\backslash x)\
(prod\ \_\ \_\ (x\backslash\ prod\ \_\ \_\ (y\backslash\ app [G', x, y])))$.

Since the two arguments have rigid heads, we start to traverse both terms
recursively by eating each lambda-abstraction. At the end of this procedure, the
remaining sub-terms are \todo{in our code this example though the eta mess
error} now $x$ and $app[G', x, y]$,

\section{HO unification in typed languages}

TODO: ho-link need the type of original term to produce a typed term in the
object language, example: f x 1 = F x => type of F = (A -> Prof) if type of f is 
(A -> nat -> Prop)

\section{Proof automation from \coq to \elpi}

TODO: representing a logic programming language into an other: compile rules
keeping higher order unification

\subsection{Dealing with FO \textit{non-syntactical} unification}

\subsection{Dealing with HO unification}

\end{document}