### 2024-04-15 paper/main.tex
section 7

\documentclass[sigconf,natbib=false,review]{acmart}
\usepackage[]{biblatex}

\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}
\addbibresource{bib.bib}

\usepackage{myTools}
\usepackage{macros}
\usepackage{enumitem}
\usepackage{intcalc}
\usepackage{nameref}

% TODO: set this fields
%\setcopyright{cc}
%\setcctype{by}
%\copyrightyear{2024}
%\acmYear{XXXX 2024}
%\acmBooktitle{YYY}
%\acmDOI{ZZZZZZZZZZZZ}

\def\githubUrl{\url{https://github.com/FissoreD/ho-unif-for-free}}

% \xspaceaddexceptions{]\}}

\def\elpi{\proglang{elpi}}
\def\coqelpi{\proglang{coq-elpi}}
\def\lambdaprolog{\proglang{$\lambda$-prolog}}
\def\coq{\proglang{coq}}

\newcommand{\library}[1]{\textit{#1}\xspace}
\def\stdpp{\library{stdpp}}
\def\iris{\library{iris}}

\newcommand*{\acronym}[1]{\texttt{#1}\xspace}

\newtheorem{invariant}{Invariant}
\crefname{invariant}{invariant}{invariants}
\crefname{equation}{property}{properties}

\def\ol{\acronym{ol}} % object language
\def\ml{\acronym{ml}} % meta language
\def\lf{\acronym{lf}} % logical framework
\def\ho{\acronym{ho}} % higher order
\def\Forall{$\forall$}

\newcommand{\appsep}{\ensuremath{\textcolor{lightgray}{\cdot}}}
\newcommand{\EqualRel}{\ensuremath{=}}
\newcommand{\nEqualRel}{\ensuremath{\new}}
\newcommand{\UnifRel}{\ensuremath{\simeq}}
\newcommand{\nUnifRel}{\ensuremath{\not\simeq}}

\newcommand{\Uo}{\texorpdfstring{\ensuremath{\UnifRel_o}\xspace}{unif\_o}}
\newcommand{\nUo}{\ensuremath{\nUnifRel_o}\xspace}
\newcommand{\Eo}{\ensuremath{\EqualRel_o}\xspace}
\newcommand{\nEo}{\ensuremath{\nEqualRel_o}\xspace}

\newcommand{\Ue}{\ensuremath{\UnifRel_\lambda}\xspace}
\newcommand{\nUe}{\ensuremath{\nUnifRel_\lambda}\xspace}
\newcommand{\Ee}{\ensuremath{\EqualRel_\lambda}\xspace}
\newcommand{\nEe}{\ensuremath{\nEqualRel_\lambda}\xspace}
\newcommand{\llambda}{\ensuremath{\mathcal{L}}\xspace}

\newcommand{\linkMacro}[1]{\ensuremath{#1}\texttt{-link}\xspace}

\newcommand{\linkbeta}{\linkMacro{\llambda}}
\newcommand{\linketa} {\linkMacro{\eta}}

\newcommand{\Fo}{\texorpdfstring{\ensuremath{\mathcal{F}_{\!o}\xspace}}{Fo}} % space non va
\newcommand{\Ho}{\texorpdfstring{\ensuremath{\mathcal{H}_o}\xspace}{Ho}}

\newcommand*{\eqtau}{\ensuremath{\mathrel{\overset{\mathrm{\tau}}{=}}}}

\newcommand{\linketaM}[3]{\ensuremath{#1 \vdash #2 =_\eta #3}}
\newcommand{\linkbetaM}[3]{\ensuremath{#1 \vdash #2 =_{\llambda} #3}}
\newcommand{\substCell}[3]{\ensuremath{#1 \vdash #2 = #3}}
\newcommand{\mapping}[3]{\ensuremath{#1 \mapsto #2^#3}}

\newcommand{\lhs}{lhs\xspace}
\newcommand{\rhs}{rhs\xspace}

\newcommand{\linkStore}{\texorpdfstring{\ensuremath{\mathbb{L}}\xspace}{L}}
\newcommand{\mapStore}{\texorpdfstring{\ensuremath{\mathbb{M}}\xspace}{M}}
\newcommand{\foUnifPb}{\ensuremath{\mathbb{P}}\xspace}
\newcommand{\hoUnifPb}{\ensuremath{\mathbb{T}}\xspace}

\include{pb_printer}

\begin{document}

% \title{HO unification from object language to meta language}
\title{Higher-Order unification for free!}
\subtitle{Reusing the meta-language unification for the object language}

\author{Davide Fissore}
\email{davide.fissore@inria.fr}
\affiliation{%
  \institution{Université Côte d'Azur, Inria}
  % \city{Nice}
  \country{France}}

\author{Enrico Tassi}
\email{enrico.tassi@inria.fr}
\affiliation{%
  \institution{Université Côte d'Azur, Inria}
  % \city{Nice}
  \country{France}}

\begin{abstract}
  Specifying and implementing a proof system from scratch requires significant effort.
  Logical Frameworks and Higher Order Logic Programming Languages provide
  dedicated, high-level Meta Languages (ML) to facilitate this task in two
  key ways: 1) variable binding and substitution are for free when ML binders
  represent object logic ones; 2) proof construction, and even proof search, are
  greatly simplified by leveraging the unification procedure provided by the ML.
  Notable examples of ML are Elf~\cite{elf}, Twelf~\cite{twelf},
  $\lambda$Prolog~\cite{miller_nadathur_2012} and
  Isabelle~\cite{10.1007/978-3-540-71067-7_7}
  which have been utilized to implement various formal systems such as
  First Order Logic~\cite{felty88cade},
  Set Theory~\cite{10.1007/BF00881873},
  Higher Order Logic~\cite{books/sp/NipkowPW02}, and even the Calculus of
  Constructions~\cite{felty93lics}.

  The object logic we are interested in is
  Coq's Calculus of Inductive Constructions (CIC)\cite{Coq-refman}. We aim to
  develop a higher-order unification-based 
  proof search procedure for it using the ML
  Elpi~\cite{dunchev15lpar}, a dialect of $\lambda$Prolog.
  Elpi's equational theory includes
  $\eta\beta$ equivalence and features a
  higher-order unification procedure \Ue restricted to the pattern
  fragment~\cite{miller92jsc}.
  Elpi offers an encoding of CIC suitable
  for meta-programming~\cite{tassi:hal-01637063,tassi:hal-01897468,gregoire:hal-03800154,newtc}
  but restricts \Ue to roughly first-order unification problems only.
  We refer to this basic encoding as \Fo.

  In this paper we propose a more well-behaved encoding called \Ho,
  and show how to translate unification problems from \Fo{}
  to corresponding ones in \Ho.
  Consequently, we derive \Uo{}, the higher-order unification procedure of \Fo{}
    that honours $\eta\beta$-equivalence (for CIC functions), addresses
  problems within the pattern fragment, and allows
  for the use of heuristics to deal with problems outside the
  pattern fragment. Moreover, as \Uo{} delegates most of the work to \Ue,
  it can be used to efficiently simulate a logic program in \Fo{} by
  taking advantage of
  unification-related optimizations of the ML, such as clause indexing.

\end{abstract}

\keywords{Logic Programming, Meta-Programming, Higher-Order Unification}

\maketitle

\section{Introduction}
\label{sec:intro}

Meta languages such as Elf~\cite{elf}, Twelf~\cite{twelf},
$\lambda$Prolog~\cite{miller_nadathur_2012}, and
Isabelle~\cite{10.1007/978-3-540-71067-7_7}
have been utilized to specify various
logics~\cite{felty88cade,books/sp/NipkowPW02,10.1007/BF00881873,felty93lics}.
The use of these meta languages facilitates this task in two
key ways. The first and most well-know one is that variable binding and
substitution come for free. %, exploited in all works mentioned above.
The second one is that these meta languages come equipped with some form
of unification, a cornerstone of proof construction and proof search.

% exploited only
%in the notable case of of Higher Order Logic~\cite{books/sp/NipkowPW02}:
%the meta language Isabelle is such a good match for HOL that it could used to
%implement an interactive proof system for the object logic, in addition to
%it specification.\\
The object logic we are interested in is Coq's~\cite{Coq-refman}
Calculus of Inductive Constructions (CIC), and we aim to implement a
form of proof search known as type-class~\cite{wadler89,sozeau08} resolution.
Type-class solvers are unification based on proof search procedures
reminiscent of Prolog, which back-chain lemmas taken
from a database of ``type-class instances''. Given this
analogy with Logic Programming we want to leverage the
Elpi~\cite{tassi:hal-01637063} meta-programming language,
a dialect of $\lambda$Prolog, already used to extend
Coq in various ways~\cite{tassi:hal-01637063,tassi:hal-01897468,gregoire:hal-03800154,newtc}.
In this paper, we focus on one aspect of
this work, precisely \emph{how to reuse the higher-order unification procedure
of the meta language in order to simulate a higher-order logic program
for the object language}.

We take as an example the \coqIn{Decision} and \coqIn{Finite} type classes
from the Stdpp~\cite{JUNG_KREBBERS_JOURDAN_BIZJAK_BIRKEDAL_DREYER_2018}
library. The class  \coqIn{Decision}
identifies predicates equipped with a decision procedure, while
\coqIn{Finite} identifies types whose inhabitants can be enumerated in a (finite) list.
The following three type-class instances state that:
1) the type of natural numbers smaller than \coqIn{n}, called \coqIn{fin n},
is finite;
2) the predicate \coqIn{nfact n nf}, relating a natural number
\coqIn{n} to the number of its prime factors \coqIn{nf}, is decidable;
3) the universal closure of a predicate has a decision procedure if
its domain is finite and if the predicate is decidable.

\begin{coqcode}
Instance fin_fin: ~$\forall$~n, Finite (fin n).             (* r1 *)
Instance nfact_dec: ~$\forall$~n nf, Decision (nfact n nf). (* r2 *)
Instance forall_dec: ~$\forall$~A P, Finite A ~$\to$~            (* r3 *)
  ~$\forall$~x:A, Decision (P x) ~$\to$~ Decision (~$\forall$~x:A, P x).
\end{coqcode}

\noindent Given this database, a type-class solver is expected to
prove the following statement automatically:

\begin{coqcode}
  Decision (~$\forall$~x: fin 7, nfact x 3)                   (* g *)
\end{coqcode}

\noindent
The proof found by the solver back-chains on rule 3 (the only rule
about the $\forall$ quantifier), and then solves the premises with
rules 1 and 2 respectively.
Note that rule 3 features a second-order parameter \coqIn{P} that represents
a function of type \coqIn{A ~$\to$~ Prop} (a predicate over \coqIn{A}).
The solver has to infer a value for \coqIn{P} by unifying the conclusion
of rule 3 with the goal, and in particular, it has to solve the unification
problem \coqIn{P x = nfact x 3}. This higher-order problem falls in the so-called 
pattern-fragment \llambda~\cite{miller92jsc} and admits a unique
solution $\customlabel{solution:intro}{\rho}$ that assigns
the term \elpiIn{~$\lambda$~x.nfact x 3} to \coqIn{P}.

In order to implement such a search in Elpi, we shall describe the encoding
of CIC terms and then the encoding of instances as rules.
Elpi comes equipped with
an Higher Order Abstract Syntax (HOAS~\cite{10.1145/53990.54010}) datatype of CIC
terms, called \elpiIn{tm}, that includes (among others) the following
constructors:

\begin{elpicode}
type lam  tm -> (tm -> tm) -> tm.     % lambda abstraction
type app  list tm -> tm.              % n-ary application
type all  tm -> (tm -> tm) -> tm.     % forall quantifier
type con  string -> tm.               % constants
\end{elpicode}

\noindent
Following the standard syntax of $\lambda$Prolog~\cite{miller_nadathur_2012},
the meta-level binding of a variable \elpiIn{x} in an expression
\elpiIn{e} is written as <<\elpiIn{x\ e}>>, while square brackets delimit a
list of terms separated by comma. For example, the term
<<\coqIn{~$\forall$~y:t, nfact y 3}>> is encoded as follows:

\begin{elpicode}
all (con"t") y\ app [con"nfact", y, con"3"]
\end{elpicode}

\noindent
We now illustrate the encoding of the three instances above as higher-order
logic programming rules: capital letters denote rule
parameters; \elpiIn{:-} separates the rule's head from the premises, and
\elpiIn{pi w\ p} introduces a fresh nominal constant \elpiIn{w}
for the premise \elpiIn{p}.

\begin{elpicode}
finite   (app [con"fin", N]).                         ~\customlabel{clause:r1}{(r1)}~
decision (app [con"nfact", N, NF]).                   ~\customlabel{clause:r2}{(r2)}~
decision (all A x\ app [P, x]) :- finite A,           ~\customlabel{clause:r3}{(r3)}~
  pi w\ decision (app [P, w]).
\end{elpicode}

\noindent
Unfortunately this intuitive encoding of rule \ref{clause:r3} does not work
since it uses the predicate \coqIn{P} as a first order term: for the meta
language its type is \elpiIn{tm}. If we try to back-chain the rule
\ref{clause:r3} on the encoding of the goal \ref{goal:g} given below

\begin{elpicode}
decision (all (app [con"fin", con"7"]) x\              ~\customlabel{goal:g}{(g)}~
  app [con"nfact", x, con"3"]).
\end{elpicode}

\noindent
we obtain an unsolvable unification problem \ref{problem:p}:
the two lists of terms have different lengths!
%The root cause is that
%\ref{problem:p} is an higher order in CIC, but becomes
%first order in the meta language due to the ``naive'' encoding.

\begin{elpicode}
app [con"nfact", x, con"3"] = app [P, x]               ~\customlabel{problem:p}{(p)}~
\end{elpicode}

\noindent
In this paper we study a more sophisticated encoding of CIC terms and rules
that, on a first approximation, would reshape \ref{clause:r3} as follows:

\begin{elpicode}
decision (all A x\ Pm x) :- link Pm P A, finite A,    ~\customlabel{clause:r3a}{(r3')}~
  pi x\ decision (app [P, x]).
\end{elpicode}

\noindent
Since \elpiIn{Pm} is a higher-order unification variable
of type \elpiIn{tm -> tm},
with \elpiIn{x}
in its scope, the unification problem \ref{problem:pa}
admits one solution:

\begin{elpicode}
app [con"nfact", x, con"3"] = Pm x                    ~\customlabel{problem:pa}{(p')}~
Pm = x\ app [con"nfact", x, con"3"]                   ~\customlabel{solution:pm}{(\sigma)}~
\end{elpicode}

\noindent
Once the head of rule \ref{clause:r3a} unifies with the goal \ref{goal:g},
the premise <<\elpiIn{link Pm A P}>> brings the assignment \ref{solution:pm}
back to the domain \elpiIn{tm} of Coq terms, obtaining the expected solution
\ref{solution:intro}:

\begin{elpicode}
P = lam A x\ app [con"nfact", x, con"3"]
\end{elpicode}

\noindent
This simple example is sufficient to show that the encoding we seek
is not trivial and does not only concern the head of rules, but the entire sequence
of unification problems that constitute the execution of a logic program.
In fact
the solution for \elpiIn{P} above generates a
(Coq) $\beta$-redex in the second premise (the predicate
under the \elpiIn{pi w\ }\hspace{-0.4em}).
% We show below the premise before and
% after the instantiation of \elpiIn{P}:

% \begin{elpicode}
% decision (app [                   P                  , w])
% decision (app [lam A (a\ app [con"nfact", a, con"3"]), w])
% \end{elpicode}

% \noindent
In turn, this redex prevents rule \ref{clause:r2} from backchaining properly since
the following unification problem has no solution:

\begin{elpicode}
app [ lam A (a\ app [con"nfact", a, con"3"]) , x] =
app [ con"nfact"                             , N, NF]
\end{elpicode}

\noindent
The root cause of the problems we outlined in this example
is a subtle mismatch between the equational theories of the meta language
and the object language, which in turn makes the 
unification procedures of the meta language weak.
The equational theory of the meta language Elpi encompasses
$\eta\beta$-equivalence and its unification procedure can solve higher-order
problems in the pattern fragment. Although 
the equational theory of CIC is much richer, for efficiency and predictability
reasons, automatic proof search procedures typically employ a unification
procedure that only captures a $\eta\beta$-equivalence and only operates
in \llambda. The similarity is striking, but one needs to exercise some caution in order to
simulate a logic program in CIC using the unification of Elpi.

\paragraph{Contributions}
In this paper we identify a minimal language \Fo{} in which the problems
sketched in the introduction can be formally described.
We detail an encoding of a logic program in \Fo{} to a strongly related
logic program in \Ho (the language of the meta-language) and we show that
the higher-order unification procedure of the meta language \Ue{} can be
efficiently used to simulate a higher-order unification procedure \Uo for
the object language that features $\eta\beta$-conversion. We show how \Uo
can be extended with heuristics to deal with problems outside the pattern
fragment.\\
Section~\ref{sec:problem-statement} formally states the problem and gives the
intuition behind our solution; \cref{sec:simulation} sets up a basic
simulation of first-order logic programs, \cref{sec:llam} and \cref{sec:eta}
extend it to higher-order logic programs in the pattern fragment
while \cref{sec:beta} goes beyond the pattern fragment.
Section \ref{sec:implementation} discusses the implementation in Elpi.\\
The $\lambda$Prolog code discussed in the paper can be accessed at the
address \githubUrl.

\section{Problem statement and solution} %%%%%%%%%%%%%%%%%%%%%%
\label{sec:problem-statement}

\newcommand{\specunif}[3]{
  \{#3_1, #3_2\} \subseteq \llambda \Rightarrow %
    \exists \rho, %
      \rho #3_1 #1 \rho #3_2  %
        \Leftrightarrow #3_1 #2 #3_2 \mapsto \rho' \subseteq \rho
}

\newcommand{\unifcorrectX}[4]{
  % \forall \rho #3_1 #3_2, %
    % \{#3_1, #3_2\} \subseteq \llambda \Rightarrow %
   #4 \Rightarrow
      #3_1 #2 #3_2 \mapsto \rho
        \Rightarrow
          \rho #3_1 #1 \rho #3_2  %
}
\newcommand{\unifcorrect}[3]{\unifcorrectX{#1}{#2}{#3}{ \{#3_1, #3_2\} \subseteq \llambda}}

\newcommand{\unifcompleteX}[4]{
  % \forall #3_1 #3_2, %
    % \{#3_1, #3_2\} \subseteq \llambda \Rightarrow %
    #4 \Rightarrow
      % \forall \rho, %
        \rho #3_1 #1 \rho #3_2  %
          \Rightarrow \exists \rho', #3_1 #2 #3_2 \mapsto \rho' \land \rho' \subseteq \rho
}
\newcommand{\unifcomplete}[3]{\unifcompleteX{#1}{#2}{#3}{\{#3_1, #3_2\} \subseteq \llambda}}

\newcommand{\maybeeta}{\texorpdfstring{\ensuremath{\Diamond\eta}\xspace}{maybeeta}}
\newcommand{\maybebeta}{\texorpdfstring{\ensuremath{\Diamond\beta}\xspace}{maybebeta}}
\newcommand{\notllambda}{\texorpdfstring{\ensuremath{\Diamond\llambda}\xspace}{maybellam}}
%\newcommand{\notllambda}{\ensuremath{\Diamond\beta}\xspace}

Even if we encountered the problem working on CIC, we devise
a minimal setting to ease its study. In this setting, we have
a \Fo{} language (for first order) with a rich equational
theory and a \Ho{} meta language with a simpler one.

\subsection{Preliminaries: \Fo{} and \Ho{}}

To reason about unification, we provide a description of the
\Fo{} and \Ho languages where unification variables
are first-class terms, i.e. they have a concrete syntax as
shown in \cref{code:common-terms}.
% We keep these languages
%minimal, for example, we omit the \elpiIn{all} quantifier of CIC we used
%in the example in Section~\ref{sec:intro} together with the type notation of
%terms carried by the \elpiIn{lam} constructor.
%
{
\setlength{\abovecaptionskip}{0pt}
\setlength{\belowcaptionskip}{-13pt}
\begin{figure}[b] %[H]
  \begin{tabular}{ll}
  \begin{minipage}{0.21\textwidth}
   \inputrawelpicode{code/fo_tm}
  \end{minipage}
  &
  \begin{minipage}{0.24\textwidth}
   \inputrawelpicode{code/ho_tm}
  \end{minipage}
  \end{tabular}\vspace{4pt}
  \caption{The \Fo{} and \Ho languages}\vspace{0.3em}
  \label{code:common-terms}
  \Description[code:common-terms]{code:common-terms}
\end{figure}
}
Unification variables
in \Fo{} (\elpiIn{fuva} term constructor) have no explicit scope:
the arguments of a higher-order variable are given via the \elpiIn{fapp}
constructor. For example the term <<\coqIn{P x}>> is represented as
<<\elpiIn{fapp [fuva N, x]}>>, where \elpiIn{N} is the memory address
of \elpiIn{P} and \elpiIn{x} is a bound variable.\\
In \Ho, the representation of <<\coqIn{P x}>> is instead <<\elpiIn{uva N [x]}>>,
since unification variables are higher-order and come equipped with an
explicit scope.
%\todo{use faddr for Fo}
%
%faddr and addr are memory addresses, the details are given in bla,
%here they are just unique identifiers for unif variables and their types
%do now allow to mistake one for the other.
%\todo{cleanup}

\paragraph{Notational conventions}

When we write \Ho terms outside code blocks we follow the
usual $\lambda$-calculus notation, reserving $f, g, a, b$ for constants,
$x, y, z$ for bound variables and $X, Y, Z, F, G, H$ for unification variables.
However, we need to
distinguish between the ``application'' of a unification variable
to its scope and the application of a term to a list of arguments.
We write the scope of unification variables in subscript
while we use juxtaposition for regular application.
Here are few examples:\\
\vspace{4pt}
{
\setlength{\tabcolsep}{1em}
\begin{tabular}{ll}
  $f\appsep a$                  & \elpiIn{app [con "f", con "a"]}\\
  $\lambda x.\lambda y.F_{x y}$ & \elpiIn{lam x\ lam y\ uva F [x, y]} \\
  $\lambda x.F_{x} \appsep a$   & \elpiIn{lam x\ app [uva F [x], con "a"]} \\
  $\lambda x.F_{x} \appsep x$   & \elpiIn{lam x\ app [uva F [x], x]} \\
\end{tabular}
}
\vspace{4pt}

\noindent
When it is clear from the context, we shall use the same syntax for \Fo{} terms
(although we never subscripts unification variables).\\
We use $s$, $s_1$, \ldots for terms in \Fo{} and $t$, $t_1$, \ldots for
terms in \Ho{}.

\subsection{Equational theories an unification}

In order to specify unification, we need to
define the equational theory and
substitution (unification-variable assignment).

\subsubsection{Term equality: \Eo and \Ee}
For both languages, we extend the equational theory
over ground terms to the full language by adding the reflexivity for
unification variables (a variable is equal to itself).

The first four rules are common to both equalities
and define the usual congruence over terms. Since
we use an HOAS encoding, they also capture $\alpha$-equivalence.
In addition to that, \Eo has rules for $\eta$ and $\beta$-equivalence.

\input{code/feq}

\begin{elpicode}
  ~\PYG{k+kd}{type} \PYG{n+nf}{(\Ee)} \PYG{k+kt}{tm -> tm -> o}~.
  con C ~\Ee~fcon C.
  app A ~\Ee~fapp B :- forall2 ~(\Ee)~ A B.
  lam F ~\Ee~flam G :- pi x\ x ~\Ee~x => F x ~\Ee~G x.
  uva N A ~\Ee~fuva N B :- forall2 ~(\Ee)~ A B.
\end{elpicode}

\noindent
The main point in showing these equality tests is to remark how
weaker \Ee is, and to identify the four rules that need special
treatment in the implementation of \Uo.
For brevity, we omit the code of \elpiIn{beta}:
it is sufficient to know that <<\elpiIn{beta F L R}>> computes in \elpiIn{R} the
weak head normal form of <<\elpiIn{app [F|L]}>>.
Note that the symbol \elpiIn{|} separates the head of a list from the tail.
\todo{explain forall2}

\paragraph{Substitution: $\rho s$ and $\sigma t$}

We write $\sigma = \{~ X \mapsto t ~\}$ for the substitution that assigns
the term $t$ to the variable $X$.
We write $\sigma t$ for the application of
the substitution to a term $t$, and $\sigma X = \{~ \sigma t ~|~ t \in X ~\}$ when
$X$ is a set of terms.
We write $\sigma \subseteq \sigma'$ when $\sigma$ is more
general than $\sigma'$.
% The domain of a substitution is the set of unification variables for which
% it provides an assignment.
% We write $\sigma \cup \sigma'$ to denote the concatenation of
% two substitutions whose domains are disjoint.
We shall use $\rho$ for \Fo{} substitutions,
and $\sigma$ for the \Ho ones.
For brevity, in this section, we consider
the substitution for \Fo{} and \Ho{} identical.
We defer to \cref{sec:grounwork} a more precise description
pointing out their differences.

\paragraph{Term unification: \Uo vs. \Ue}

% Although we provide an implementation of the meta-language unification \Ue
% in the supplementary material (that we used for testing purposes) we only
% describe its signature here.
\Ho{}'s unification signature is:

\input{code/ue_type}

\noindent
We write 
$\sigma t_1 \Ue \sigma t_2 \mapsto \sigma'$ when
$\sigma t_1$ and $\sigma t_2$ unify with substitution $\sigma'$.
Note that $\sigma'$ is a refined (i.e. extended) version of $\sigma$; this is
reflected by the signature above that relates two substitutions.
We write $t_1 \Ue t_2 \mapsto \sigma'$ when
the initial substitution $\sigma$ is empty.
% Note that if $\sigma t_1 \Ue \sigma t_2 \mapsto \sigma'$ then
% the domains of $\sigma$ and $\sigma'$ are disjoint.
We write \llambda as the set of terms that are in the pattern-fragment, i.e.
every higher-order variable is applied to a list of distinct names.

The meta language of choice is expected to provide 
an implementation of \Ue that satisfies
the following properties:
%~\ref{prop:correct-ml} and~\ref{prop:complete-ml}.
\begin{gather}
  \unifcorrect{\Ee}{\Ue}{t} \label{prop:correct-ml}\\
  \unifcomplete{\Ee}{\Ue}{t}\label{prop:complete-ml}
\end{gather}

\noindent
Even if we provide an implementation of the object-language unification
\Uo{} in \cref{sec:founif}, our real goal is the simulation of an entire
logic program.

\subsection{The problem: logic-program simulation}
We represent a logic program \emph{run} in \Fo{} as
a sequence of \emph{steps} of length $\mathcal{N}$.
At each step, $p$ we unify two terms, $\foUnifPb_{p_l}$ and
$\foUnifPb_{p_r}$, taken from the list of all unification
problems \foUnifPb.
% \footnote{If the same rule is used multiple times in a run we
% consider as many copies as needed of the terms composing the
% rules, with fresh unification variables each time. Also, we do not
% model backtracking directly, but rather the fact that a step in a run may
% fail, and in that case the entire run is said to fail. A regular logic program
% that succeeds is modeled as a many runs, with possibly overlapping starts,
% where at least one run does not end in a failure. }
The composition of these steps starting from the
empty substitution $\rho_0$ produces the final
substitution $\rho_\mathcal{N}$, which is the result of the
logic program execution.
%
\newcommand{\C}[4]{\ensuremath{\langle #1 \rangle}\mapsto(#2,#3,#4)}
\newcommand{\D}[4]{\ensuremath{\langle #1,#2,#3 \rangle^{-1}\mapsto #4}}
\newcommand{\progress}{\ensuremath{\mathrm{progress}}\xspace}
\newcommand{\fstep}{\ensuremath{\mathrm{fstep}}\xspace}
\newcommand{\hstep}{\ensuremath{\mathrm{hstep}}\xspace}
\newcommand{\frun}{\ensuremath{\mathrm{frun}}\xspace}
\newcommand{\hrun}{\ensuremath{\mathrm{hrun}}\xspace}
\newcommand{\stepF}[4]{\ensuremath{\fstep(#1,#2,#3) \mapsto #4}}
\newcommand{\stepFD}[4]{%
\ensuremath{#3 #1_{#2_l} \Uo #3 #1_{#2_r} \mapsto #4}}
\newcommand{\stepH}[6]{\ensuremath{\hstep(#1,#2,#3,#4) \mapsto (#5, #6)}}
\newcommand{\stepHD}[6]{\ensuremath{%
#3 #1_{#2_l} \Ue #3 #1_{#2_r} \mapsto #4 \land \progress(#6,#4) \mapsto (#6',#5)}}
\newcommand{\runF}[3]{\ensuremath{\frun(#1,#2) \mapsto #3_{#2}}}
\newcommand{\runFD}[2]{\ensuremath{%
\bigwedge_{p = 1}^{#2} \stepF{#1}{p}{\rho_{p-1}}{\rho_{p}}}}
\newcommand{\runH}[3]{\ensuremath{\hrun(#1,#2) \mapsto #3_{#2}}}
\newcommand{\runHD}[3]{\ensuremath{%
\bigwedge_{p = 1}^{#2} \stepH{#1}{p}{\sigma_{p-1}}{#3_{p-1}}{\sigma_{p}}{#3_p}}}
\newcommand{\deff}{\ensuremath{\stackrel{{\scriptscriptstyle def}}{=\!=\!\!\!=}}}
%
$$
\begin{array}{l}
\stepF{\foUnifPb}{p}{\rho}{\rho'}
\deff
\stepFD{\foUnifPb}{p}{\rho}{\rho'}\vspace{2pt}\\
\runF{\foUnifPb}{\mathcal{N}}{\rho}
\deff
\runFD{\foUnifPb}{\mathcal{N}}
\end{array}
$$

\noindent
In order to simulate a \Fo{} logic program in \Ho{}, we compile
each \Fo{} term $s$ in \foUnifPb to a \Ho{} term $t$.
We write this translation as $\C{s}{t}{m}{l}$. The implementation of the compiler
is detailed in \cref{sec:simulation,sec:eta,sec:beta}, here we just point
out that it additionally produces a variable mapping $m$ and a list of links $l$.
The variable map connects unification variables in \Ho to variables
in \Fo{} and is used to ``decompile'' the assignment,
$\D{\sigma}{m}{l}{\rho}$. Links are an accessory piece of information whose
description is deferred to \cref{sec:nutshell}.
We write $\hoUnifPb_p = \{~ \hoUnifPb_{p_l}, \hoUnifPb_{p_r} ~\}$
and $s \in \foUnifPb \Leftrightarrow \exists p, s \in \foUnifPb_p$.

We simulate each run in \Fo{} with a run in \Ho as follows:
%
%Note that $\sigma_0$ is the empty substitution.
$$
\begin{array}{l}
\stepH{\hoUnifPb}{p}{\sigma}{\linkStore}{\sigma''}{\linkStore'} \deff\vspace{2pt}\\
  \qquad\stepHD{\hoUnifPb}{p}{\sigma}{\sigma'}{\sigma''}{\linkStore}\vspace{2pt}\\
  \runH{\foUnifPb}{\mathcal{N}}{\rho} \deff \vspace{2pt}\\
  \qquad \hoUnifPb \times \mapStore \times \linkStore_0 = \{ (t,m,l) | s \in \foUnifPb, \C{s}{t}{m}{l} \}\vspace{2pt}\\
  \qquad \runHD{\hoUnifPb}{\mathcal{N}}{\linkStore}\vspace{2pt}\\
  \qquad \D{\sigma_{\mathcal{N}}}{\mapStore}{\linkStore_{\mathcal{N}}}{\rho_{\mathcal{N}}}
\end{array}
$$

\noindent
By analogy with \foUnifPb, we write $\hoUnifPb_{p_l}$ and $\hoUnifPb_{p_r}$
for the two \Ho{} terms being unified at step $p$, and we write $\hoUnifPb_p$
for the set $\{~ \hoUnifPb_{p_l}, \hoUnifPb_{p_r} ~\}$.\\
\hstep{} is made of two sub-steps: a call to the meta language
unification and a check for \progress{} on the set of links, that intuitively
will compensate for the weaker equational theory honored by \Ue.
\hrun{} compiles all terms in \foUnifPb{}, then executes each step, and
finally decompiles the solution.
We claim:

\begin{proposition}[Simulation]\label{prop:simulation}
$\forall \foUnifPb, \forall \mathcal{N},$ if $~\foUnifPb \subseteq \llambda$
$$
  \runF{\foUnifPb}{\mathcal{N}}{\rho}
  \Leftrightarrow
  \runH{\foUnifPb}{\mathcal{N}}{\rho}
$$
\end{proposition}

\noindent
That is, the two executions give the same result if all terms in
\foUnifPb are in the pattern fragment. Moreover:

\begin{proposition}[Simulation fidelity]\label{prop:fidelity}
In the context of$\;$ \hrun, if $~\foUnifPb \subseteq \llambda$ we have that
$\forall p \in 1 \ldots \mathcal{N},$
$$
\stepF{\foUnifPb}{p}{\rho_{p-1}}{\rho_p}
\Leftrightarrow
\stepH{\hoUnifPb}{p}{\sigma_{p-1}}{\linkStore_{p-1}}{\sigma_p}{\linkStore_p}
$$
\end{proposition}
\noindent
In particular, this property guarantees that a \emph{failure} in the \Fo{} run
is matched by a failure in \Ho{} \emph{at the same step}. We consider this
property very important from a practical point of view since it guarantees
that the execution traces are strongly related, and in turn, this enables a user
to debug a logic program in \Fo{} by looking at its execution trace in
\Ho{}.

We also claim that \hrun handles terms outside \llambda in the following sense:

\begin{proposition}[Fidelity recovery]\label{prop:fidelity-recovery} 
In the context of \hrun, if 
$\rho_{p-1} \foUnifPb_{p} \in \llambda$ 
(even if $\;\foUnifPb_{p} \not\in \llambda$)
then
$$
\stepF{\foUnifPb}{p}{\rho_{p-1}}{\rho_p} \Leftrightarrow
\stepH{\hoUnifPb}{p}{\sigma_{p-1}}{\linkStore_{p-1}}{\sigma_{p}}{\linkStore_p}
$$
\end{proposition}

In other words, if the two terms involved in a step
re-enter \llambda, then \hstep and \fstep are again related, even if
$\foUnifPb \not\subseteq \llambda$ and hence \cref{prop:fidelity} does not apply.
Indeed, the main difference between \cref{prop:fidelity} and \cref{prop:fidelity-recovery}
is that the assumption of the former is purely static, it can be checked upfront. 
When this assumption is not satisfied, one can still simulate a logic program and
have guarantees of fidelity if, at run time, decidability of higher-order
unification is restored.

This property has practical relevance since in many logic programming
implementations, including Elpi, the order in which unification problems
are tackled does matter.
The simplest example is the sequence $F \UnifRel \lambda x.a$ and
$F \appsep a \UnifRel a$: the second problem is not in \llambda and has two
unifiers, namely $\sigma_1 = \{~ F \mapsto \lambda x.x ~\}$ and
$\sigma_2 = \{~ F \mapsto \lambda x.a ~\}$. The first problem picks $\sigma_2$,
making the second problem re-enter \llambda.

\paragraph{Backtracking} We omit it from our model of a logic program's execution
since it plays a very minor role, orthogonal to higher-order unification.
We point out that each \emph{run} corresponds to a (proof search) branch in the
logic program that either fails at some point, or succeeds. A computation that
succeeds by backtracking, exploring multiple branches, could be
modeled as a set of runs with (possibly non-empty) common prefixes.

\subsection{The solution (in a nutshell)}
\label{sec:nutshell}
A term $s$ is compiled to a term $t$ where every
``problematic'' sub term $p$ is replaced by a fresh unification variable $h$
with an accessory \emph{link} that represents a suspended unification problem
$h \Ue p$. As a result \Ue is ``well behaved'' on $t$, in the sense that
it does not contradict \Eo as it would otherwise do on the
``problematic'' sub-terms.\\
We now define ``problematic'' and ``well behaved'' more formally
We use the $\Diamond$ symbol since it stands for ``possibly'' in modal logic
and all problematic terms are characterized by some ``uncertainty''.

\begin{definition}[\maybebeta]\label{def:maybebeta}
  \maybebeta is the set of terms of the form $X \appsep x_1 \ldots x_n$
  such that $x_1 \ldots x_n$ are distinct names (of bound variables).
\end{definition}

\noindent
An example of a \maybebeta{} term is the application $F \appsep x$.
This term is problematic since the application node of
its syntax tree cannot be used to justify a
unification failure, i.e. by properly instantiating $F$ the term
head constructor may become a $\lambda$, or a constant, or remain an application.

\begin{definition}[\maybeeta]\label{def:maybeeta}
  \maybeeta is the set of terms $s$ such that $\exists \rho, \rho s$
  is an eta expansion.
\end{definition}

\noindent
An example of a term $s$ in \maybeeta{} is
$\lambda x.\lambda y.F \appsep y \appsep x$
since the substitution
$\rho = \{ F \mapsto \lambda a.\lambda b.f \appsep b \appsep a\}$
makes $\rho s = \lambda x.\lambda y.f \appsep x \appsep y$,
which is the eta long form of $f$. This term is problematic since
its leading $\lambda$ abstraction cannot justify a
unification failure against a constant $f$.

\begin{definition}[\notllambda]\label{def:notllambda}
  \notllambda is the set of terms of the form $X \appsep t_1 \ldots t_n$
  such that $t_1 \ldots t_n$ are not distinct names.
\end{definition}

\noindent
These terms are problematic for the very same reason terms in \maybebeta are,
but they cannot be handled directly by the unification of the meta language, which
is only required to handle terms in \llambda. Still, there exists a
substitution $\rho$ such that $\rho s \in \llambda$.

% An example of $t$ in \notllambda{} is $F \appsep a$ for a constant $a$.
% Note however tha
% an oracle could provide an assignment $\rho = \{ F \mapsto \lambda x.x\}$
% that makes the resulting term fall back in \llambda.
% \todo{said before}

\newcommand{\subterm}[1]{\ensuremath{\mathcal{P}(#1)}}
We write $\subterm{t}$ the set of sub-terms of $t$, and
%\begin{definition}[Subterms \subterm{t}] The set of sub terms of $t$ is the
%  largest set $\subterm{t}$ that can be obtained by the following rules.
%$$
%\begin{array}{l}
%t \in \subterm{t}\\
%t = f t_1\ldots t_n \Rightarrow \subterm{t_i} \subseteq \subterm{t} \land f \in \subterm{t}\\
%t = \lambda x.t' \Rightarrow \subterm{t'} \subseteq \subterm{t}\\
%\end{array}
%$$
%\end{definition}
%\noindent
we write $\subterm{X} = \bigcup_{t\in X} \subterm{t}$ when $X$ is a set of terms.

\newcommand{\wellb}{\ensuremath{\mathcal{W}}\xspace}
\begin{definition}[Well behaved set]
Given a set of terms $X \subseteq \Ho{}$,
$$
\wellb(X) \Leftrightarrow \forall t \in \subterm{X},
t \not\in (\maybebeta{} ~\cup~ \maybeeta{} ~\cup~ \notllambda{})
$$
\end{definition}

\noindent
We write $\wellb(t)$ as a short for $\wellb(\{t\})$.
We claim our compiler validates the following property:

\begin{proposition}[\wellb-enforcing]\label{prop:w-enforcing}
  Given two terms $s_1$ and $s_2$, if $\exists \rho, \rho s_1 \Eo \rho s_2$,
  then 
  $$
    \C{s_i}{t_i}{m_i}{l_i} ~\mathrm{for}~ i \in \{ 1,2 \} \Rightarrow
    t_1 \Ue t_2 \mapsto \sigma \label{prop:compilation-w}
  $$
\end{proposition}

\noindent
In other words the compiler outputs terms in \wellb, even if its
input is not.
Note that the property holds for any substitution. $\rho$ could be given by an
oracle and/or not necessarily be a most general one: in \wellb \Ue{} simply
does not contradict \Eo.

\begin{proposition}[\wellb{}-preservation]\label{prop:w-preservation}
$\forall \hoUnifPb, \forall \linkStore, \forall p, \forall \sigma, \forall \sigma'$
$$
\begin{array}{l}
\wellb(\sigma\hoUnifPb) \land
  \sigma\hoUnifPb_{p_l} \Ue \sigma\hoUnifPb_{p_r} \mapsto {\sigma'}
  \Rightarrow \wellb(\sigma' \hoUnifPb)\\
\wellb(\sigma\hoUnifPb) \land
  \progress(\linkStore,\sigma) \mapsto (\_,\sigma')
  \Rightarrow \wellb(\sigma' \hoUnifPb)
\end{array}
$$
\end{proposition}

\noindent
Proposition \ref{prop:w-preservation} is key to proving \cref{prop:simulation,prop:fidelity}.
Informally, it says that the problematic terms moved on the side by the compiler
are not reintroduced by \hstep, hence \Ue{} can continue to operate properly.
In \cref{sec:simulation,sec:eta,sec:beta}
we describe how the compiler recognizes terms in \maybebeta, \maybeeta and
\notllambda and how \progress takes care of them preserving \wellb
and ensuring \cref{prop:simulation,prop:fidelity,prop:fidelity-recovery}.

% Note that proposition \ref{prop:w-preservation} does not hold for \hrun as a whole
% since decompilation can introduce (actually restore) problematic terms.

\section{Basic compilation and simulation}
\label{sec:simulation}

\subsection{Memory map (\mapStore) and substitution (\texorpdfstring{$\rho$ and $\sigma$}{rho and sigma})}
\label{sec:grounwork}

% The predicate to test this condition is called \elpiIn{pattern-fragment}:

% \input{code/pattern_fragment}

Unification variables are identified by a (unique) memory address.
The memory and its associated operations are described below:

\input{code/mem_types}

\noindent
If a memory cell is \elpiIn{none}, then the corresponding unification variable
is not set. \elpiIn{assign} sets an unset cell to the given value, while
\elpiIn{new} finds the first unused address and sets it to \elpiIn{none}.

Since each \Ho unification variable occurs together with a scope,
its assignment needs to be abstracted over it to enable the
instantiation of the same assignment to different scopes.
This is expressed by the \elpiIn{inctx} container, and in particular
its \elpiIn{abs} binding constructor.

\input{code/ho_subst}

\noindent
A solution to a \Fo{} variable is a plain term, that is \elpiIn{fsubst}
is an abbreviation for \elpiIn{mem fm}.
%We call \elpiIn{fsubst} the memory of \Fo{}, while we call \elpiIn{subst}
%the one of \Ho.

The compiler establishes a mapping between variables of the two languages.

\input{code/comp_base_types}

\noindent
Each \elpiIn{hvariable} is stored in the mapping together with
its arity (a number) so that the code of \ref{clause:malloc} below can preserve:

\begin{invariant}[Unification-variable arity]\label{inv:uvaarity}
  Each variable \elpiIn{A}
  in \Ho has a (unique) arity \elpiIn{N} and each occurrence
  \elpiIn{(uva A L)} is such that \elpiIn{L} has length \elpiIn{N}.
\end{invariant}

\input{code/alloc_mapping}

\noindent
When a single \elpiIn{fvariable} occurs multiple times with different numbers
of arguments, the compiler generates multiple mappings for it, on a first
approximation, and then ensures the mapping are bijective by introducing
\linketa; this detail is discussed in section \ref{sec:invariant1}.

%Applying the substitution corresponds to dereferencing a term with respect to
%the memory.
%  To ease the comparison we split \Fo{} dereferencing into a
% \elpiIn{fder} step and a \elpiIn{napp} one. The former step replaces references
% to memory cells that are set with their values, and has a corresponding
% operation in \Ho, namely \elpiIn{deref}. On the contrary \elpiIn{napp}
% has no corresponding operation in \Ho, and only ensures that
% terms of the form <<\elpiIn{fapp[fapp L1|L2]}>> are replaced by
% <<\elpiIn{fapp L3}>> where \elpiIn{L3} is the concatenation of \elpiIn{L1}
% and \elpiIn{L2}. The reasons for this asymmetry is
% that an \elpiIn{fapp} node with a flexible head is always mapped
% to a \elpiIn{uva} (as per \cref{sec:simulation,sec:beta}),
% preventing nested applications to materialize.
% \input{code/fderef}
It is worth examining the code of \elpiIn{deref}, which
applies the substitution to a \Ho{} term. Notice how assignments are moved
to the current scope, i.e. the \elpiIn{abs}-bound variables are renamed
with the names in the scope of the unification variable occurrence.

\input{code/deref}

\noindent
Note that \elpiIn{move} strongly relies on invariant \ref{inv:uvaarity}: the length
of the arguments of all occurrences of a unification variable is the same. Hence,
they have the same simple type for the meta-level, and therefore the number of
\elpiIn{abs} nodes in the assignment matches that length.
This guarantees that \elpiIn{move} never fails.

\input{code/move}

We write $\sigma = \{~ A_{xy} \mapsto y ~\}$ for the assignment
<<\elpiIn{abs x\abs y\y}>> and $\sigma = \{~ A \mapsto \lambda x.\lambda y.y ~\}$
for <<\elpiIn{lam x\lam y\y}>>.

\subsection{Links (\linkStore)}

\noindent
As mentioned in section~\ref{sec:nutshell}, the compiler
replaces terms in \maybeeta, \maybebeta, and \notllambda with fresh
variables linked to the problematic terms. Terms in \maybebeta do not
need a link since \Ho{} variables faithfully represent
the problematic term thanks to their scope.

\input{code/comp_links}

\noindent
The right-hand side of a link, the problematic term, can occur under binders.
To accommodate this situation, the compiler wraps \elpiIn{baselink} using
the \elpiIn{inctx} container (see~\ref{data:inctx} also used for \elpiIn{subst}).

\begin{invariant}[Link left hand side]\label{inv:linklhs}
  The left-hand side of a suspended link
  is a variable.
\end{invariant}

\noindent
New links are suspended by construction.
If the left-hand side is assigned during a step, then 
the link is considered for progress and possibly eliminated.
This is discussed in \cref{sec:eta} and \cref{sec:beta}.

When detailing examples, we represent links as equations between two
terms under a context.
The equality sign is subscripted with the
kind of \elpiIn{baselink}. For example $\linkbetaM{x}{A_x}{F_x~a}$ corresponds to:
\begin{elpicode}
abs x\ val (link-llam (uva A [x]) (app[uva F [x],con "a"]))
\end{elpicode}

% The first three rules unify terms with same rigid heads, and
% call the unification relation on the sub-terms. If $t_1$ (resp. $t_2$) is an
% assigned variables, $t_1$ is dereferenced to $t_1'$ (resp. $t_2'$) and the
% unification is called between $t_1'$ and $t_2$ (resp. $t_1$ and $t_2'$). If both
% terms are unification variables, we test that their arguments are in the pattern
% fragment, we allocate a new variable $w$ in $\rho_1$ such that $w$ is the
% pruning of the arguments of $t_1$ and $t_2$, we assign both $t_1$ and $t_2$ to
% $w$ and return the new mapping $\rho_2$ containing all the new variable
% assignment. Finally, if only one of the two terms is an unification variable
% $v$, after having verified that $v$ does not occur in the other term $t$, we
% bind $v$ to $t$ and return the new substitution mapping.

% \old

% A key property needed in unification is being able to verify if two terms are
% equal wrt a given equational theory. This relation allow to compare terms under
% a certain substitution mapping, so that any time a variable $v$ is assigned in a
% subterm, a dereferencing of $v$ is performed. After variable dereferencing, the
% test for equality is continued on the new-created subterm.

% The base equality function over terms can be defined as follows:

% The solution we are proposing aim to overcome these unification issues by 1)
% compiling the terms $t$ and $u$ of the OL into an internal version $t'$ and $u'$
% in the ML; 2) unifying $t'$ and $u'$ at the meta level instantiating meta
% variables; 3) decompiling the meta variable into terms of the OL; 4) assigning
% the variables of the OL with the decompiled version of their corresponding meta
% variables. We claim that $t$ and $u$ unify if and only if $t'$ and $u'$ unify
% and that the substitution in the object language is the same as the one returned
% by the ML. \noteE[inline]{same or $\supseteq$ or $\subseteq$}

% In the following section we explain how we deal with term (de)compilation and
% links between unification variables.

% \section[Compilation: fo\_tm to tm]{Basic simulation of \Fo{} in \Ho{}}

% In this section we describe a basic compilation scheme that we refine
% later, in the following sections. This scheme is sufficient to implement
% a \hstep that respects $\beta$-conversion for terms in \llambda.
% The extension to $\eta\beta$-conversion is described in \cref{sec:eta} and
% the support for terms outside \llambda in \cref{sec:beta}.

\subsection{Compilation}
\label{sec:compilation}

The simple compiler described in this section serves as a base for the
extensions in \cref{sec:llam,sec:eta,sec:beta}.
Its main task is to beta normalize the term and map one syntax tree to
the other.
In order to bring back the substitution from \Ho{} to \Fo{} the compiler
builds a ``memory map'' connecting the kind of variables using routine
\ref{clause:malloc}.\noteD{non mi piace}

The signature of the \elpiIn{comp} predicate below allows for the generation of
links (suspended unification problems), which play no role in this section
but play a major role in \cref{sec:llam,sec:eta,sec:beta}.
With respect to \cref{sec:problem-statement}, the signature also allows
for updates to the substitution.

\input{code/comp_base}
\input{code/compile.tex}

\noindent
The code above uses that possibility
in order to allocate space for the variables, i.e. it sets their memory
address to \elpiIn{none} (a details not worth mentioning in the
previous sections).
\todo{explain fold6}

\input{code/comp_lam}

\noindent
In the code above, the syntax \elpiIn{pi x y\..} is syntactic sugar for
iterated \elpiIn{pi} abstraction, as in \elpiIn{pi x\ pi y\..}.

The auxiliary function \elpiIn{close-links} tests if the bound variable
\elpiIn{v} really occurs in the link. If it does, the link is wrapped into
an additional \elpiIn{abs} node binding \elpiIn{v}. In this way links generated
deep inside the compiled terms can be moved outside their original context
of binders.

\input{code/comp_close_links}

\noindent
Note that we could remove the first rule, whose sole purpose is to make
links more readable by pruning unused context entries.

\subsection{Execution}
\label{sec:execution}

A step in \Ho consists of unifying two terms and reconsidering all
links for progress. If either of these tasks fails, we consider the entire step to
fail. It is at this granularity that we can relate steps in the
two languages.

\input{code/hstep}

\noindent
Note that the infix notation \elpiIn{((A ~\Ue~B) C D)} is syntactic sugar for
\elpiIn{((~\Ue\!\!\!~) A B C D)}.

Reconsidering links is a fixpoint process because the progress of a link
can update the substitution, which may then enable another link to progress.

\input{code/progress}

\subsubsection{Progress}
In the base compilation scheme, \elpiIn{progress1} is the identity function
on both the links and the substitution, so the fixpoint trivially terminates.
Sections \ref{sec:eta} and \ref{sec:beta} add rules to \elpiIn{progress1}
and explain why the don't hinder termination.
% For brevity we omit the code
% that applies the substitution \elpiIn{S1} to all terms in \linkStore.

\subsubsection{Occur check}\label{sec:oc}
Since compilation moves problematic terms out of the sight of \Ue{},
that procedure can only perform a partial occur check. For example, the
unification problem $X \Ue f~Y$ cannot generate a cyclic substitution alone,
but should be disallowed if a $\linkStore$ contains a link like
$\linketaM{}{Y}{\lambda z.X_z}$: we don't know yet if $Y$ will feature
a lambda in head position, but we surely know it contains $X$, hence
$f~Y$ and that fails the occur check.
The procedure \elpiIn{occur-check-links} is in charge of
performing this check that is needed in order to
guarantee \fullref{prop:fidelity}.

\subsection{Substitution decompilation}\label{sec:decompilation}

Decompiling the substitution involves three steps.

First and foremost, problematic terms stored in
\linkStore have to be moved back into the game:
a suspended link must be turned into a valid assignment.
This operation is possible thanks to \fullref{inv:linklhs},
which ensures that no link causes an occur-check~(\ref{sec:oc})
and the fact that \linkStore is duplicate-free (\cref{TODO}).

The second step involves allocating new variables in the
memory of \Fo{}. This technicality is required because some
higher-order unifications may require pruning a variable. For example,
$F\appsep x\appsep y = F\appsep x\appsep z$ requires allocating a 
variable $G$ in order to express the assignment $F_{ab} \mapsto G_a$.

The final step is to decompile each assignment.
Decompiling a term is straightforward since \mapStore is a bijection.
The only complex part concerns the \elpiIn{abs} node. In our simple setting,
the \elpiIn{flam} node carries no additional information (other than the function body),
so each \elpiIn{abs} node can be trivially converted to a \elpiIn{flam} one.
However, in the case of CIC, where lambdas carry the type of the bound variable, one
must store this information somewhere. Note that this information is similar
to the arity of variables; that in CIC, unification variables have a
(function) type, and this type can be used to annotate the lambdas
needed to express their assignment.

\begin{lemma}[Compilation round trip]\label{prop:comprt} If
  \elpiIn{compile S T [] M [] _ [] _} then \elpiIn{decompile M T S}
\end{lemma}\noteD{Should also decompile links\dots}
% \begin{proof}[Proof sketch]
% Trivial since \mapStore is a bijection and 
% the terms are beta normal.
% some discussion about commit maybellam to be done later.
% \end{proof}

\subsection{Definition of \Uo{} and its properties}\label{sec:founif}

We already have all the pieces to show the code of \Ue{}.

\input{code/unif_fo}

\noindent
So far the compiler is very basic. It does not really
enforce that the terms passed to \hstep are in \wellb,
and indeed makes no use of the higher-order capabilities
of the meta language (all generated
variables have an empty scope).
Still, we can prove that \Uo is a good
``first-order'' unification algorithm if the input
already happens to be in \wellb.
Later, when the compiler will enforce \cref{prop:w-enforcing}
the proof will be adjusted to cover for the new cases.

\begin{lemma}[Properties of \Uo]\label{prop:uniffow} The following properties hold for \Uo:
\begin{gather}
  \unifcorrectX{\Ee}{\Ue}{t}{\wellb(\{t_1, t_2\})}\\
  \unifcompleteX{\Ee}{\Ue}{t}{\wellb(\{t_1, t_2\})}
\end{gather}
\end{lemma}
\begin{proof}[Proof sketch]
 In this setting, \Ee is as strong as \Eo on ground terms. What we have to show is that whenever two different \Fo{}
terms can be made equal by a substitution $\rho$ (plus the \ref{clause:beta1}
and \ref{clause:beta2} if needed), we can produce this $\rho$ by finding
a $\sigma$ via \Ue{} on the corresponding \Ho terms and by decompiling it.
If we look at the syntax of \Fo{}, terms the only interesting case is
\elpiIn{fuva X ~\Uo~s}. In this case, after compilation, we have
$Y \Ue t$ that succeeds with $\sigma = \{ Y \mapsto t\}$ and
$\sigma$ is decompiled to $\rho = \{ Y \mapsto s\}$ by \cref{prop:comprt}.
\end{proof}

\begin{theorem}[Fidelity in \wellb] \Fullref{prop:simulation} and
\fullref{prop:fidelity} hold if $\;\wellb{(\foUnifPb)}$.
\end{theorem}
\begin{proof}[Proof sketch]
Trivial since \elpiIn{progress1} is a no-op and \fstep and \hstep are the same,
and by \cref{prop:uniffow} \Ue is equivalent to \Uo.
\end{proof}

\subsection{Notational conventions}

In the following sections, we adopt this notation discuss the compiler's
output: \foUnifPb represents the input set of problems that compiled to
\hoUnifPb with memory mapping \mapStore and links \linkStore. For example:
%
\printAlll
  {{{p_1,p_2},{p_3,p_4}}}
  {{{t_1,t_2},{t_3,t_4}}}
  {{{X_1,A_1,{{x}}},{X_2,A_2,{{y}}}}}
  {{{\eta,\Gamma,a,b}}}

We index each sub-problem, sub-mapping, and sub-link with its position
starting from $1$ and counting from left to right, top to bottom.
For example, $\hoUnifPb_2$ corresponds to the \Ho problem $t_3 \Ue t_4$.
%The compiled version of each $\foUnifPb_i$ is represented by $\hoUnifPb_i$.

% \subsection{Limitations of by this basic scheme}
% \label{sec:basic-comp-limitations}
% The basic compilation scheme is not about to
% deal wit the following problem:
% \printAlll
%   {{{\lambda x y.X \appsep y \appsep x, \lambda x y.x},{\lambda x. f \appsep (X \appsep x) \appsep x,Y}}}
%   {{}}
%   {{}}
%   {{}}
% % \begin{gather}
% % \lambda x y.F \appsep y \appsep x = \lambda x y.x \label{eq:unif-eta1}\\
% % \lambda x. f \appsep (F \appsep x) \appsep x = G \label{eq:unif-eta2}
% % % \lambda x. f \appsep (F \appsep x) \appsep x = f \appsep (\lambda y.y) \label{eq:unif-eta2}
% % \end{gather}

% \noindent Note that here $X$ is used with different arities, moreover
% in the second problem the left hand side happens to be an
% eta expansion (of $f (\lambda y.y)$) only after we discover (at run time)
% that $X = \lambda x\lambda y.y$ (i.e. that $X$ discards the $x$ argument).
% Both problems are addressed in the next two sections.

\section{Handling of \maybebeta}\label{sec:llam}

In order to make \Uo higher-order, we need to
take care of terms in \maybebeta.
In the example below, we can see that the basic compilation given in 
the previous section is not able to make the \Ho unification problem succeeds.
%
% make test ONLY=7006 TEX=tex
\printAlll
  {{{\lambda x.(f\appsep (X\appsep x)\appsep a),\lambda x.(f\appsep x\appsep a)}}}
  {{{\lambda x.(f\appsep (A\appsep x)\appsep a),\lambda x.(f\appsep x\appsep a)}}}
  {{{X,A,0}}}
  {{}}
\todo{fix indentation}
The unification problem $\hoUnifPb_1$ fails while trying to unify
$A \appsep x$ and $x$, which is equivalent to \elpiIn{app [uva A [], x]}
versus \elpiIn{x}.
In order to exploit the higher-order unification algorithm of the meta language,
we need to compile the \Fo{} term $X\appsep x$ into the \Ho term $A_x$,
which is \elpiIn{uva A [x]}.

\subsection{Compilation and decompilation}

We add the following rule before rule~\ref{rule:compapp}, where
\elpiIn{pattern-fragment} is a predicate checking if a list of
terms is a list of distinct names.

\input{code/comp_base_beta}

\noindent
Note that compiling \elpiIn{Ag} cannot create new mappings nor links, since \elpiIn{Ag}
is made of bound variables, and the hypothetical rule 
\ref{rule:complamh}
loaded by \elpiIn{comp-lam}
grants this property.

% The only detail worth discussing is the fact that the procedure updates a
% substitution, rather than just crafting one as presented in
% section~\ref{sec:problem-statement}. The reason is that the algorithm folds
% over a term, updating a substitution while it traverses it.\noteE[inline]{explain better}

\paragraph{Decompilation} Since no link is created by the compilation of 
\maybebeta terms, no modification to the \elpiIn{commit-link}
is needed.

\paragraph{Progress} Similarly to decompilation, since no link is produced,
no modification to the \elpiIn{progress} predicate is needed.

\begin{definition}[$\wellb/_\beta$]
$
\wellb/_\beta(X) \Leftrightarrow \forall t \in \subterm{X}, t \not\in (\maybeeta{} ~\cup~ \notllambda{})
$
\end{definition}

\begin{lemma}[Properties of \Uo]\label{prop:uniffowb} The following properties hold for \Uo
  where
  \begin{gather}
    \unifcorrectX{\Ee}{\Ue}{t}{\wellb/_\beta(\{t_1, t_2\})}\\
    \unifcompleteX{\Ee}{\Ue}{t}{\wellb/_\beta(\{t_1, t_2\})}
  \end{gather}
\end{lemma}

  \begin{proof}[Proof sketch]
  If we look at the \Fo{} terms, there is one more interesting case,
  namely \elpiIn{fapp[fuva X|W] ~\Uo~s} when \elpiIn{W}
  are distinct names compiled to $\vec{w}$. In this case
  the \Ho problem is $Y_{\vec{w}} \Ue t$ that succeeds with
   $\sigma = \{ Y_{\vec{y}} \mapsto t[\vec{w}/\vec{y}]\}$, which in turn
   is decompiled to $\rho = \{ Y \mapsto \lambda \vec{y}.s[\vec{w}/\vec{y}]\}$.
   Thanks to \ref{clause:beta1}
   $(\lambda \vec{y}.s[\vec{w}/\vec{y}])~\vec{w} \Eo s$.
  \end{proof}

\begin{lemma}[$\wellb/_\beta$-enforcement]\label{lem:w-enforcement-maybebeta}
  Given two terms $s_1$ and $s_2$ in \maybebeta, if $\exists \rho, \rho s_1 \Eo \rho s_2$,
  then 
  $$
    \C{s_i}{t_i}{m_i}{l_i} ~\mathrm{for}~ i \in \{ 1,2 \} \Rightarrow
    t_1 \Ue t_2 \mapsto \sigma
  $$ 
\end{lemma}
\begin{proof}[Proof sketch]
  problematic terms are mapped to uva by comp,
  the problematic fapp node is gone.
  if F x = t then  F = lam x.t[]
  \todo{finish/maybe W enforcement is not needed here}
\end{proof}

\begin{theorem}[Fidelity in $\wellb/_\beta$] \Fullref{prop:simulation} and
\fullref{prop:fidelity} hold if $\;\wellb/_\beta(\foUnifPb)$
\end{theorem}
\begin{proof}[Proof sketch] thanks to \cref{lem:w-enforcement-maybebeta}
  \Ue is as powerful as \Uo in \maybebeta, as well as in \wellb by
  \cref{prop:uniffowb}.
\end{proof}

% \subsection{Example}

% OK

% \begin{elpicode}
% Terms [ flam x\ fapp[fcon"g",fapp[fuva z, x]]
%       , flam x\ fapp[fcon"g", fcon"a"] ]
% Problems = [ pr z (s z) ] % $\lambda x.g (F x) = \lambda x.g a$
% lam x\ app[con"g",uva z [x]] ~\Uo~lam x\ app[con"g", con"a"]
% link z z (s z)
% HS = [some (abs x\con"a")]
% S = [some (flam x\fcon a)]
% \end{elpicode}

% KO

% \begin{elpicode}
%   Terms [ flam x\ fapp[fcon"g",fapp[fuva z, x]]
%   , flam x\ fapp[fcon"g", fcon"a"] ]
% Problems = [ pr 0 1   % $A = \lambda x.x$
%            , pr 2 3 ] % $A a = a$
% lam x\ app[con"g",uva z [x]] ~\Uo~lam x\ app[con"g", con"a"]
% link z z (s z)
% HS = [some (abs x\con"a")]
% S = [some (flam x\fcon a)]
% lam x\ app[f, app[X, x]] = Y,
%   lam x\ x) = X.
% \end{elpicode}

% \otext{Goal: $s_1 \Uo s_2$ is compiled into $t_1 \Ue t_2$}
% \otext{What is done: uvars \elpiIn{fo_uv} of OL are replaced into uvars \elpiIn{ho_uv} of the ML}
% \otext{Each \elpiIn{fo_uv} is linked to an \elpiIn{ho_uv} of the OL}
% \otext{Example needing the compiler v0 (tra l'altro lo scope è ignorato):\\ \elpiIn{lam x\ app[con"g",app[uv 0, x]] ~\Uo~lam x\ app[con"g", c"a"]}}
% \otext{Links used to instantiate vars of elpi}
% \otext{After all links, the solution in links are compacted and given to coq}
% \otext{It is not so simple, see next sections (multi-vars, eta, beta)}

% The compilation step is meant to recover the higher-order variables of the OL,
% expressed in a first order way, by replacing them with higher-order variables in
% the ML. In particular, every time a variable of the OL is encountered in the
% original term, it is replaced with a meta variable, and if the OL variable is
% applied to a list of distinct names $L$, then this list becomes the scope of the variable.
% For all the other constructors of
% \elpiIn{tm}, the same term constructor is returned and its arguments are
% recursively compiled. The predicate in charge for term compilation is:

% \elpiIn{type comp tm -> tm -> links -> links -> subst -> subst -> o}.

% \noindent
% where, we take the term of the OL, produce the term of the ML, take a list
% of link and produce a list of new links, take a substitution and return a
% new substitution.

% In particular, due to programming constraints, we need to drag the old subst and
% return a new one extended, if needed, with the new declared meta-variables.

% The following code
% %
% \begin{elpicode}
%   kind link type.
%   type link nat -> nat -> nat -> subst.
% \end{elpicode}
% %
% \noindent
% defines a link, which is a relation between to variables indexes, the first
% being the index of a OL variable and the second being the index of a ML
% variable. The third integer\noteE[inline]{integer or nat?} is the number of term in the
% scope of the two variables, or equivalently, in a typed language, their arity.

% As an example, let's study the following unification problem (a slightly
% modified version from \cref{sec:intro}):

% \begin{elpicode}
%   lam x\ app[c"decision", app[c"nfact", x, c"3"]] ~\Uo~
%     lam x\ app [c"decision", app[uv 0, x]]
% \end{elpicode}

% \noindent
% we have the main unification problem where the nested \elpiIn{app} nodes have
% lists of different lengths making the unification to fail. The compilation of
% these terms produces a new unification problem with the following shape:

% \begin{elpicode}
%   lam x\ app[c"decision", app[c"nfact", x, c"3"]] ~\Ue~
%     lam x\ app [c"decision", uv 1 [x]]
% \end{elpicode}

% \noindent
% The main difference is the replacement of the subterm \elpiIn{app[uv 0, x]} of
% the OL with the subterm \elpiIn{uv 0 [x]}. Variable indexes are chosen by the
% ML, that is, the index \elpiIn{0} for that unification variable of the OL term
% has not the sam meaning of the index \elpiIn{0} in the ML. There exists two
% different substitution mapping, one for the OL and one for the ML and the indexes
% of variable point to the respective substitution.

% decomp che mappa abs verso lam
% \noindent
% \otext{An other example: \\
%   \elpiIn{lam x\ app[f, app[X, x]] = Y, (lam x\ x) = X.}}

% \section{Use of multivars}

% Se il termine initziale è della forma

% \begin{elpicode}
%   app[con"xxx", (lam x\ lam y\ Y y x), (lam x\ f)]
%   =
%   app[con"xxx",X,X]
% \end{elpicode}

% allora se non uso due X diverse non ho modo di recuperare il quoziente che mi manca.

% a sto punto consideriamo liste di problemi e così da eliminare sta xxx senza
% perdità di generalità (e facciamo problemi più corti, e modellizziamo anche la
% sequenza)

\section{Handling of \maybeeta}\label{sec:eta}
A term 
$\lambda x. t \appsep x$ is said to be the $\eta$-expansion of $t$ if
$x$ does not occur free in $t$, and conversely, $t$ is the $\eta$-contraction of
$\lambda x. t \appsep x$. The equational theory of \Fo{} identifies these terms,
but the current compilation scheme does not,
as shown by the following example:
%
\printAlll
  {{{\lambda x. X \appsep x, f}}}
  {{{\lambda x. A_x, f}}}
  {{{X,A,1}}}
  {{}}

\noindent
While $\lambda x.X \appsep x \Uo{} f$ does admit the solution
$\rho = \{~ X \mapsto f ~\}$, the corresponding problem in
\hoUnifPb does not:
\elpiIn{lam x\ uva A [x]} and
\elpiIn{con"f"} start with different, rigid, term constructors hence
\Ue{} fails.

In order to guarantee \cref{prop:simulation}, we detect
lambda abstractions that can disappear by $\eta$-contraction
(\cref{sec:etadetection}), and we modify the compiler so that it
generates fresh unification variables
in their place and moves the problematic term 
from \hoUnifPb to \linkStore (\cref{sec:etacomp}). The compilation
of the problem \foUnifPb above is refined to: 
%
\printAlll
  {{{\lambda x.X \appsep x,f}}}
  {{{A,f}}}
  {{{X,B,1}}}
  {{{\eta,,A,\lambda x.B_x}}}

\noindent
As per \cref{inv:linklhs} the \linketa left-hand side is a variable
while the right-hand side is a term in \maybeeta that has the following property:

\begin{invariant}[\linketa \rhs]
  The \rhs of any \linketa %in \linkStore 
  has the shape $\lambda x.t$
  and $t$ is not a lambda. 
  %where $t_x$ is a \maybeeta term and $x$ is free in $t$.
  \label{inv:link-eta-right}
\end{invariant}

Each \linketa is kept in the link store \linkStore during execution
and is activated under some conditions.
Activation is implemented by extending the \elpiIn{progress1}
predicate defined in \cref{sec:execution}.

\subsection{Detection of \maybeeta}\label{sec:etadetection}

When compiling a term $t$, we need to determine if any
subterm $s \in \subterm{t}$ that is of the form $\lambda x. r$,
where $x$ occurs in $r$, can be a $\eta$-expansion, i.e. if
there exists a substitution $\rho$ such that $\rho (\lambda x.r) \Eo s$.
The detection of lambda abstractions that can ``disappear''
is not as trivial as it may seems. Here a few examples:
%
\begin{center}
  \begin{tabular}{lll}
    %Term & Status & Evidence \\\hline
    $\lambda x. f \appsep (A \appsep x)$ & $\in\maybeeta$ & $\rho = \{~ A \mapsto \lambda x.x ~\}$ \\
    $\lambda x. f \appsep (A \appsep x) \appsep x$ & $\in\maybeeta$ & $\rho = \{~ A \mapsto \lambda x.a ~\}$\\
    $\lambda x. f \appsep x \appsep (A \appsep x)$ & $\not\in\maybeeta$ &\\
    $\lambda x. \lambda y. f \appsep (A \appsep x) \appsep (B \appsep y \appsep x)$ & $\in\maybeeta$ & $\rho = \{~ A \mapsto \lambda x.x,~ B \mapsto \lambda y.\lambda x.y ~\}$
  \end{tabular}
\end{center}
\vspace{4pt}

% In the examples above, the first term is a \maybeeta since $A_x$ can
% reduce to $x$ by setting $A_x = \lambda x.x$, 
% the second one is not a \maybeeta since it exists no substitution
% for $A_x$ such that $A_x$ reduces to $x$ and $x$ is not free in the subterm $f \appsep x$.
\noindent
The first two examples are easy, and show how a unification variable can expose
or erase a variable in its scope, turning the resulting term into an $\eta$-expansion or not.\\
The third example shows that when a variable occurs outside the scope of a unification
variable, it cannot be erased and can hence prevent a term from being an $\eta$-expansion.\\
The last example shows the recursive nature of the check we need to implement.
The term starts with a spine of two lambdas, hence the whole term
is in \maybeeta iff the inner term $\lambda y.f\appsep (A \appsep x) \appsep (B \appsep y \appsep x)$
is in \maybeeta itself. If it is, it could $\eta$-contract to
$f \appsep (A \appsep x)$ making $\lambda x.f \appsep (A \appsep x)$ a potential
$\eta$-expansion.\\
We can now define more formally how \maybeeta terms are detected together with
its auxiliary functions:

\newcommand{\reduceto}{\emph{may-contract-to}}
\begin{definition}[\reduceto]
  A $\beta$-normal term $s$ \reduceto{} a name $x$ if there exists a
  substitution $\rho$ such that $\rho s \Eo{} x$.
\end{definition}

\begin{lemma}\label{lem:reduceto}
A $\beta$-normal term $s = \lambda x_1 \ldots x_n.t$
%, where $x, x_1\ldots x_n$ can occur in $t$, 
\reduceto{} $x$ only if one of the following three conditions holds:
\begin{enumerate}
  \item $n = 0$ and $t = x$;
  \item $t$ is the application of $x$ to a list
     of terms $l$ and each $l_i$ \reduceto{} $x_i$
     (e.g. $\lambda x_1 \ldots x_n.x \appsep x_1 \ldots x_n \Eo{} x$) ;
  \item $t$ is a unification variable with scope
    $W$, and for any $v \in \{ x, x_1 \ldots x_n \}$,
    there exists a $w_i \in W$, such that $w_i$ \reduceto{} $v$
    (if $n = 0$ this is equivalent to $x \in W$).
\end{enumerate}
\end{lemma}
\begin{proof}[Proof sketch]
Since our terms are in $\beta$-normal form there is only
one rule that can play a role (namely \ref{clause:eta1}),
hence if the term $s$ is not exactly $x$ (case 1)
it can only be an $\eta$-expansion of $x$, or a unification
variable that can be assigned to $x$, or a combination of both.
If $s$ begins with a lambda, then the lambda can only disappear by $\eta$
contraction. In that case, the term $t$ is under the spine of binders
$x_1\ldots x_n$, $t$ can either be $x$ applied to terms that can
\reduceto{} these variables (case 2), or a unification variable
that can be assigned to that application (case 3).
\end{proof}

% \noindent
% Note that this condition does not require the term to be in \llambda.
% \noteE[inline]{Is this relevant}

\newcommand{\occursrigid}{\emph{occurs-rigidly}\xspace}
\newcommand{\occurrigid}{\emph{occur-rigidly}\xspace}
\begin{definition}[\occursrigid]\label{def:occrigid}
  A name $x$ \occursrigid{} in a $\beta$-normal term $t$, if $\forall \rho, x \in
  \subterm{\rho t}$
\end{definition}

In other words, $x$ \occursrigid in $t$ if it occurs in $t$
outside of the scope of a unification variable $X$; otherwise, an instantiation
of $X$ can make $x$ disappears from $t$.
Moreover, note that $\eta$-contracting $t$ cannot make $x$ disappear, since
$x$ is not a locally bound variable inside $t$.

We can now describe the implementation of \maybeeta detection:

\newcommand{\testmaybeeta}{\emph{maybe-eta}\xspace}
\begin{definition}[\testmaybeeta]\label{def:testmaybeeta}
  Given a $\beta$-normal term
  $s = \lambda x_1 \ldots x_n.t$, \testmaybeeta{} $s$ holds if any
  of the following holds:
  \begin{enumerate}
    \item $t$ is a constant or a name applied to the arguments
      $l_1 \ldots l_m$ such that 
      $m \geq n$ and for every $i$ such that $m - n < i \leq m$
      the term  $l_i$
      \reduceto{} $x_i$, and
      no $x_i$ \occursrigid{} in $l_1 \ldots l_{m-n}$;
    \item $t$ is a
      unification variable with scope $W$ and
      for each $x_i$ there exists a $w_j \in W$ such that $w_j$
      \reduceto{} $x_i$.
  \end{enumerate}
\end{definition}
\begin{lemma}[\maybeeta detection]\label{lem:maybeeta}
  If $t$ is a $\beta$-normal term
  and
  $t \in \maybeeta$ then \testmaybeeta{} $t$ holds.
\end{lemma}
\begin{proof}[Proof sketch]
Follows from \cref{def:occrigid} and \cref{lem:reduceto}
\end{proof}

\noindent
Remark that the converse of \cref{lem:maybeeta} does not hold: 
there exists a term $t$ satisfying the criteria (1) of
\cref{def:testmaybeeta} that is not in $\maybeeta$, i.e.
there exists no substitution $\rho$ such that $\rho t$ is an
$\eta$-expansion. A simple counter example is
$\lambda x. f \appsep (A \appsep x) \appsep (A \appsep x)$
since $x$ does not \occurrigid{} in the first argument
of $f$,
and the second argument of $f$ \reduceto{} $x$.
In other words $A \appsep x$
may either use or discard $x$, but our analysis does not
take into account that \emph{the same
term} cannot have two contrasting behaviors.

As we will see in the rest of this section, this is not a problem
since it does not break
\cref{prop:simulation} nor \cref{prop:fidelity}.
% A term in \maybeeta{} is compiled to a
% unification variable and a link (see \cref{sec:etacomp}):
% the link makes progress (see \cref{sec:etaprogress})
% in the same step in which the
% variable is instantiated, and that compensates
% for this coarse analysis.

% The implementation we propose for the \maybeeta relation is given below.

% \input{code/maybe_eta}

% Here a complex maybeeta example
% \begin{gather}
%   T = \lambda y. f \appsep A_{xy} \appsep (B \appsep a \appsep (\lambda z.y\appsep C_z)) \appsep D_x
% \end{gather}

\subsection{Compilation and decompilation}\label{sec:etacomp}

% Thanks to the \elpiIn{maybe-eta} predicate, we can detect ``$\eta$-problematic''
% terms and, consequently replace them with fresh \Ho unification variables at
% compilation time. The code below illustrate how this relation is used to for
% term compilation.

\paragraph{Compilation}
The following rule is inserted just before rule~\ref{rule:complam} from the code in
\cref{sec:compilation}.

\input{code/comp_eta}

\noindent
Whenever \elpiIn{flam F} is detected to be in
\maybeeta it is compiled to \elpiIn{lam F1} and replaced by the fresh
variable \elpiIn{A}. This variable sees all the names free in
\elpiIn{lam F1}. The variable is connected to \elpiIn{lam F1} via a \linketa.
Invariant~\fullref{inv:linklhs} holds for this link, moreover:

\begin{corollary}
  The \rhs of any \linketa has exactly one lambda abstraction, hence
  the rule above respects \fullref{inv:link-eta-right}.
  \label{cor:rhs-eta-onelamb}
\end{corollary}

\begin{proof}[Proof sketch]
  By contradiction, suppose that the rule above is applied and that
  the \rhs of the link is $\lambda x.\lambda y.t$, where $x$ and $y$ occur in $t$.
  If $\testmaybeeta{}~\lambda y.t$ holds then the recursive call to
  \elpiIn{comp} (made by \elpiIn{comp-lam}) must have put a fresh variable
  in its place, so this case is impossible.
  Otherwise, if $\testmaybeeta{}~\lambda y.t$ does not hold, then also
  $\testmaybeeta{}~\lambda x.\lambda y.t$ does not hold either, contradicting
  the assumption that the rule was applied.
\end{proof}

\paragraph{Decompilation}
The decompilation of a \linketa is performed by unifying
the \lhs with the \rhs. Note that this unification never fails, since \lhs is a
flexible term not appearing in any other \linketa
(by \cref{def:progressetadedup}).

\subsection{Progress}\label{sec:etaprogress}

\linketa{}s are meant to delay the unification of ``problematic'' terms until
we know for sure if the term has to be $\eta$-contracted or not.

\newcommand{\progressetaleft}{\emph{$\eta$-progress-\lhs}\xspace}
\begin{definition}[\progressetaleft]\label{def:progressetaleft}
A link \linketaM{\Gamma}{X}{T} is removed from \linkStore when
$X$ becomes rigid. Let $y\in\Gamma$, there are two cases:
\begin{enumerate}
  \item if $X = a$ or $X = y$ or $X = f \appsep a_1\ldots a_n$
    we unify the $\eta$-expansion of $X$ with $T$, that is we run
    $\lambda x.X \appsep x\Ue{} T$
    \item if $X = \lambda x.t$ we run $X \Ue{} T$.
\end{enumerate}
\end{definition}

\newcommand{\progressetaright}{\emph{$\eta$-progress-\rhs}\xspace}
\begin{definition}[\progressetaright]\label{def:progressetaright} A link
\linketaM{\Gamma}{X}{T} is removed from \linkStore when either 1)
$\testmaybeeta~T$ does not hold (anymore) or 2) by $\eta$-contracting $T$ to
$T'$ where $T'$ does not start with the \elpiIn{lam} constructor. In the
first case, $X$ is unified with $T$, and in the second one, $X$ is unified with
$T'$ (under the context $\Gamma$).
\end{definition}

There is a third case in which a link is removed from \linkStore, namely
when the \lhs is assigned to a variable that is the \lhs of another
\linketa.

\newcommand{\progressetadedup}{\emph{$\eta$-progress-deduplicate}\xspace}
\begin{definition}[\progressetadedup]\label{def:progressetadedup}
  A link \linketaM{\Gamma}{X_{\vec{s}}}{T} is removed from \linkStore when
  another link \linketaM{\Delta}{X_{\vec{r}}}{T'} is in  \linkStore.
  By \cref{inv:uvaarity} the length of $\vec{s}$ and $\vec{r}$ is the same;
  hence we can move the term $T'$ from $\Delta$ to $\Gamma$ by renaming its
  bound variables, i.e. $T'' = T'[\vec{r}/\vec{s}]$.
  We then run $T \Ue{} T''$ (under the context $\Gamma$).
\end{definition}

\begin{lemma}
  Let $\lambda x.t$ the \rhs of a \linketa, then $\wellb(t)$.
  \label{lemma:unif-eta-aux}
\end{lemma}

\begin{proof}[Proof sketch]
  By construction, every ``problematic'' term in \Fo{} is replaced with a variable
  in the corresponding \Ho term. Therefore, $t$ is \wellb.
\end{proof}

% \noteD[inline]{Below the proof of \cref{prop:w-preservation}, ho usato 3 lemmi ausiliari,
% forse si può compattare in una prova più piccola?}
\begin{lemma}
  Given a \linketa $l$, the unification done by \progressetaleft is between
  terms in $\wellb$  
  \label{lemma:unif-eta-1}
\end{lemma}

\begin{proof}[Proof sketch]
  Let $\sigma$ be the substitution, which is  $\wellb(\sigma)$ (by
  \todo{is w sigmaa defined?}
  \cref{prop:w-preservation}). $\lhs \in \sigma$, therefore $\wellb(\lhs)$. By
  \progressetaleft, if 1) the \lhs is a name, a constant or an
  application, then, $\lambda x.\lhs \appsep x$ is unified with the \rhs. By
  \cref{inv:link-eta-right} and \cref{lemma:unif-eta-aux},
  $\rhs = \lambda x. t$ and $\wellb(t)$.
  Otherwise, 2) the \lhs has \elpiIn{lam} as functor.
  \todo{what does as a functor mean?}
  In both cases, unification is
  performed between terms in \wellb.
\end{proof}

\begin{lemma}
  Given a \linketa $l$, the unification done by \progressetaright is between
  terms in \wellb.
  \label{lemma:unif-eta-2}
\end{lemma}

\begin{proof}[Proof sketch]
  \lhs is variable, and, by \cref{def:progressetaright}, \rhs is either no more
  a \maybeeta, i.e. \rhs is not a $\eta$-expansion and, so, $\wellb(\rhs)$,
  otherwise, \rhs can reduce to a term which cannot be a $\eta$-expansion, and,
  so, $\wellb(\rhs)$. In both cases, the unification between \rhs and \lhs
  is done between terms that are in \wellb.
\end{proof}

\begin{lemma}
  Given a \linketa $l$, the unification done by \progressetadedup is between
  terms in \wellb.
  \label{lemma:unif-eta-3}
\end{lemma}

\begin{proof}
  We unify the \rhs{} of two \linketa. Both have the
  shape $\lambda x.t$ and by \cref{lemma:unif-eta-aux}, $\wellb(t)$.
\end{proof}

\begin{lemma}
  The introduction of \linketa guarantees \fullref{prop:w-enforcing}
  %\label{lemma:unif-wellb}
\end{lemma}

\begin{proof}[Proof sketch]
  By \cref{def:testmaybeeta}.\todo{shall we move the proof here?}
\end{proof}

\begin{lemma}
  The progress of \linketa guarantees \fullref{prop:w-preservation}
  \label{lemma:unif-wellb}
\end{lemma}

\begin{proof}[Proof sketch]
  By \cref{lemma:unif-eta-1,lemma:unif-eta-2,lemma:unif-eta-3}, every
  unification performed by the activation of a \linketa is done between
  terms in \wellb, therefore, the substitution remains \wellb.
\end{proof}

\begin{lemma}
  \elpiIn{progress} terminates.
  \label{lemma:prog-eta-terminates}
\end{lemma}

\begin{proof}[Proof sketch]
  Rules \cref{def:progressetaleft,def:progressetaright} and
  \cref{def:progressetadedup} remove one link from \linkStore, hence they
  cannot be applied indefinitely.
  Moreover each rule only relies on terminating operations such as \Ue,
  $\eta$-contraction, $\eta$-expansion, relocation (a recursive copy of a
  finite term).
  % The addition of rules for \elpiIn{progres1} complicates the function
  % \elpiIn{progress}. We can note, however, that they do not prevent the
  % termination of \elpiIn{progress}. 1) If a link is activated it is removed from
  % \linkStore and the recursive call to \elpiIn{progress} will have a smaller
  % list of links to recurse on. Moreover, link activation only runs terminating
  % instructions (such as unification). 2) If a link is deduplicated, the
  % termination of \elpiIn{progress} is still guaranteed since again we reduce
  % \linkStore and the instructions run by link deduplications are all
  % terminating. 3) If a link is neither activated nor deduplicated, i.e. it
  % remains suspended, then \linkStore remains unchanged like the substitution;
  % therefore, \elpiIn{if (L = L1, S1 = S2)} succeeds and \elpiIn{progress}
  % terminates.
\end{proof}

\begin{theorem}[Fidelity in \maybeeta]
  % F x = z, F x y = z y
  Given a list of unification problems \foUnifPb, if \noteD{riprendere sec 3.1}
  $\forall t, t\in\subterm{\foUnifPb} \Rightarrow t \notin \notllambda$ and if the 
  memory map is bijective, then
  the introduction 
  of \linketa guarantees \fullref{prop:fidelity}.
  \footnote{This premise is fulfilled in \cref{sec:invariant1}}
  \label{lemma:fidelity-maybeeta}
\end{theorem}

\begin{proof}[Proof sketch]
  \progressetaleft and \progressetadedup activate a \linketa when, in the
  original unification problem, a \maybeeta term is unified with respectively a
  well-behaved term or another \maybeeta term. In both cases, the links trigger
  a unification which succeeds iff the same unification in \Fo{} succeeds,
  \todo{Fix: show that 5.7 an 5.8 compensate for missing eta rule}
  guaranteeing \cref{prop:fidelity}. \progressetaright never fails, in fact,
  this progression refines a variable to a rigid term and plays no role in 
  \cref{prop:fidelity}.
  % If 1) a \linketa is activated by \progressetaleft, then unification is done
  % between a \maybeeta term $t_1$ and a term $t_2$ (with $\wellb(t_2)$). This
  % activation performs a unification which succeeds iff the original problem in
  % \Fo{} succeeds. If 2) a \linketa is activated by \progressetadedup, then the
  % unification is done between two \maybeeta terms, and again this unification
  % succeeds iff it succeeds in \Fo{}. Finally, if 3) a \linketa is activated by
  % \progressetaright, the unification, done between a variable and a term,
  % always succeeds, this is what we expect to guarantee fidelity, since we
  % are essentially removing a hole added by the compilation in the place of a
  % \maybeeta subterm. In all the cases fidelity is respected.
\end{proof}

\paragraph{Example of \progressetaleft}

The example at the beginning of \cref{sec:eta}, once
$\sigma = \{~ A \mapsto f ~\}$, triggers \progressetaleft since the link
becomes \linketaM{}{f}{\lambda x.B_{x}} and the \lhs is a constant.
This rule runs $\lambda x.f \appsep x \Ue{} \lambda x.B_{x}$,
resulting in $\sigma = \{~ A \mapsto f ~;~ B_{x} \mapsto f ~\}$.
By the definition of \linketa decompilation,
decompilation generates $\rho = \{~ X \mapsto \lambda x. f \appsep x ~\}$, since:
$X$ is mapped to $B_x$ and $B_x$ is decompiled into $\lambda x.f \appsep x$.
% \todo{explain why decompile does so and which eta?}

\paragraph{Example of \progressetadedup}

A very basic example of \linketa deduplication, is given below:
% make test ONLY=7002 TEX=tex
\printAlll
  {{{\lambda x.(X\appsep x),\lambda x.(Y\appsep x)}}}
  {{{A,C}}}
  {{{X,B,1},
    {Y,D,1}}}
  {{{\eta,,A,\lambda x.B_{x}},
    {\eta,,C,\lambda x.D_{x}}}}

\noindent
The result of $A \Ue{} C$ is that the two \linketa share the same \lhs.
By unifying the two \rhs we get
$\sigma = \{ A~ \mapsto C, B \mapsto D ~\}$.
In turn, given the map \mapStore, this second assignment is decompiled to
$\rho = \{~ X \mapsto Y ~\}$ as expected.

We give an example for rule
\progressetaright in the next section.

%\section{Enforcing inv.: \NoCaseChange{\nameref{inv:uvaarity}}}
\section{Making \mapStore a bijection}
\label{sec:invariant1}

In \cref{sec:grounwork}, we introduced the definition of memory map \mapStore.
This map allows to decompile the \Ho substitution to the object language.
Given the higher order nature of CIC the same unification
variable $X$ is used multiple times with different arities.
%
% make test ONLY=7001 TEX=tex
\printAlll
  {{{\lambda x.\lambda y.(X\appsep y\appsep x),\lambda x.\lambda y.x},
    {\lambda x.(f\appsep (X\appsep x)\appsep x),Y}}}
  {{{A,\lambda x.\lambda y.x},
    {D,F}}}
  {{{X,E,1},
    {Y,F,0},
    {X,C,2}}}
  {{{\eta,,D,\lambda x.(f\appsep E_{x}\appsep x)},
    {\eta,,A,\lambda x.B_{x}},
    {\eta,x,B_{x},\lambda y.C_{y x}}}}

In the unification problems \foUnifPb above, we see that $X$ is used with arity $2$ in
$\foUnifPb_1$ and with arity $1$ in $\foUnifPb_2$. 
In order to preserve \fullref{inv:uvaarity} the compiler
did generate two entries in \mapStore for $X$, and in turn this
breaks \fullref{prop:fidelity} since two incompatible assignements to
the variables $X$ maps to, namely $E$ and $C$, would not be detected.
We remedy by post-processing the compiler's output by adding
additional \linketa to ensure the folllowgin property:

\begin{proposition}[\mapStore is a bijection]
  After compilation, for each \Fo{} variable $X$ in \foUnifPb{} and for each \Ho{}
  variable $A$ in \hoUnifPb{} there is exactly one entry $X \mapsto A^n$
  (for any arity $n$).
\end{proposition}

\noindent
Note that the execution in \Ho{} may require allocating new variables as
explained in~\cref{sec:decompilation}. The property above does not
concern these.

The core procedure of this post-processing step is \emph{align-arity} that
is iterated by \elpiIn{map-deduplication}:

\newcommand{\alignarity}{\emph{align-arity}}
\begin{definition}[\alignarity] Given two mappings
  $m_1 : X \mapsto A^m$ and
  $m_2 : X \mapsto C^n$ where $m < n$ and $d = n - m$,
  $\alignarity{}~ m_1 ~ m_2$ generates the following $d$ links, one
  for each $i$ such that $0 \leq i < d$,
\[%
  \linketaM{x_0 \ldots x_{m+i}}
           {B^i_{x_0 \ldots x_{m+i}}}
           {\lambda x_{m+i+1}.B^{i+1}_{x_0 \ldots x_{m+i+1}}}
\]%
where $B^i$ is a fresh variable of arity $m+i$, and $B^0 = A$ as well as $B^d = C$.
\end{definition}

The intuition is that we $\eta$-expand the occurrence of the variable
with lower arity to match the higher arity. Since each \linketa can
add exactly one lambda, we need as many links as the difference between the
two arities.

\newcommand{\mapdeduplication}{\emph{map-deduplication}\xspace}

\begin{definition}[\mapdeduplication]
  Forall mappings $m_1, m_2\in \mapStore$ such that 
  $m_1 : X \mapsto A^m$ and
  $m_2 : X \mapsto C^n$ and $m < n$
  we remove $m_1$ from \mapStore and
  add to \linkStore the result of $\alignarity{}~m_1~m_2$.
\end{definition}

\begin{theorem}[Fidelity with \mapdeduplication]
  Given a list of unification problems \foUnifPb, such that 
  $\forall t, t\in\subterm{\foUnifPb} \Rightarrow \wellb(t) \lor t \in \maybeeta$,
  if $\;\foUnifPb$ contains the same \Fo{} variable used at different arities,
  then \mapdeduplication guarantees \fullref{prop:fidelity}
\end{theorem}

\begin{proof}[Proof sketch]
  By the definition of \mapdeduplication, any two occurrencies of the same \Fo{} 
  variables $X_1, X_2$ with different arities are compiled to two
  variables $A^n$ and $C^m$ related by a chain of \linketa.
  We show that whenever one of the two is instantiated with a rigid term the 
  the chain becomes shorter and hence, eventually, the
  other variable is instantiated too.
  Without loss of generality we assume $m < n$ and $A$ is instantiated.
  The corresponding \linketa is triggered by \progressetaleft instantiating
  the next variable in the chain of links.
\end{proof}
\todo{Not sure I fullfy fixed the proof, but the old oen was not ok}

\noindent
Note that shortening the chain of links allows unification to fail if
$X_1$ and $X_2$ are unified with different terms.

If we look back the example give at the beginning of this section, we can
deduplicate $\mapping{X}{E}{1}, \mapping{X}{C}{2}$ by removing the first mapping
and adding the auxiliary \linketa: \linketaM{x}{E_{x}}{\lambda y.C_{x y}}.
After deduplication the compiler output is as follows:
% make test ONLY=7001 TEX=tex
\printAlll
  {{{\lambda x.\lambda y.(X\appsep y\appsep x),\lambda x.\lambda y.x},
    {\lambda x.(f\appsep (X\appsep x)\appsep x),Y}}}
  {{{A,\lambda x.\lambda y.x},
    {D,F}}}
  {{{Y,F,0},
    {X,C,2}}}
  {{{\eta,x,E_{x},\lambda y.C_{x y}},
    {\eta,,D,\lambda x.(f\appsep E_{x}\appsep x)},
    {\eta,,A,\lambda x.B_{x}},
    {\eta,x,B_{x},\lambda y.C_{y x}}}}

\noindent
The resolution of $\hoUnifPb_1$ assigns $\lambda x.\lambda y.x$ to $A$
that in turn triggers $\linkStore_3$ and then $\linkStore_4$ by \progressetaleft,
one per $\lambda$ in the solution of $A$.
The unification variable $C_{yx}$ is therefore unified with $x$
(the second variable of its scope).
As a result $\linkStore_1$ becomes \linketaM{x}{E_x}{\lambda y.y},
and the \rhs is no more in \maybeeta so \progressetaright fires and
$E_x$ is unified with $\lambda y.y$, resulting in
$\sigma = \{~ A \mapsto \lambda x.\lambda y.x ~;~ B_x \mapsto \lambda y.x ~;~ C_{yx} \mapsto x ~;~ E_x \mapsto \lambda y.y~\}$.\\
The remaining step $\hoUnifPb_2$ identifies $D$ with $F$,
hence the resulting \Fo{} substitution is
\todo{verificare, la vecchia soluzione era sballata}
$\rho = \{X \mapsto \lambda x.\lambda y.y ~;~ Y \mapsto \lambda x.(f\appsep (\lambda y.y))\appsep x\}$.

\section{Handling of \notllambda}\label{sec:beta}

Even if we do not requrie \Ue to handle problems outside of \llambda it is
worth handling terms in \notllambda since, in practice, these terms often
re-enter \llambda at runtime, as observed in~\cite{Nadathur2001} for example,
or to accommodate for the introduction of heuristics.

In the following example problem $\foUnifPb_2$ is $X \appsep a \Uo a$  that
admits two different substitutions: $\rho_1 = \{X \mapsto \lambda x.x\}$
and $\rho_2 = \{X \mapsto \lambda x.a\}$. The unification algorithm alone
cannot chose, since no solution is more general than the other,
although the first problem solve this dilemma at run-time making the second
problem well behaved:
%
\printAlll
  {{{X,\lambda x.a},
    {(X\appsep a),a}}}
  {{{A,\lambda x.a},
    {(A\appsep a),a}}}
  {{{X,A,0}}}
  {{}}

% In this section we suppose the unification of the object language between two
% terms $t_1$ and $t_2$ to fail each time at least one of the between $t_1$ or
% $t_2$ is outside $\llambda$. This means for instance that $X \nUo Y \appsep Z$ and 
% $X\appsep Y \nUo X\appsep Y$.

% On
% the other hand, we also point out that the \maybeeta detection spot out
% potential $\eta$-expansion for terms that are not in \llambda. For example,
% $\lambda x.F \appsep G_x$ is considered as \maybeeta, since we have the
% application of term whose argument can reduce to $x$.

% \noteE[inline]{IIRC, second order unification, where variables can stand for functions, is semi-decidable, indeed Huet's algorith enumerates all solutions. Third order, functions that take functions, is undecidable, proved by Dowek. IMO here the problem is that there is no unique solution, (semi)decidability is a secondary point. This paragraph needs fixing.}
% \noteD[inline]{I've rewritten it, it is clearer?}
% In general, unification between \notllambda terms admits more then one
% solution and committing one of them in the substitution does not guarantee
% \cref{prop:complete-ml}. For instance, $X \appsep a \Uo a$ 
% admits two different substitutions: $\rho_1 = \{X \mapsto \lambda x.x\}$
% and $\rho_2 = \{X \mapsto \lambda \_.a\}$. Prefer one over the other may break
% future unifications.

% Given a list of unification problems, $\foUnifPb_1\dots
% \foUnifPb_n$ with $\foUnifPb_n$ in \notllambda, it is often the case that the resolution of
% $\bigwedge_{i=0}^{n-1}\foUnifPb_i$ gives a partial substitution $\rho$, such
% that $\rho\foUnifPb_n$ falls again in \llambda.
% %
% % make test ONLY=7003 TEX=TEX + deactivate beta tex
% \printAlll
%   {{{X,\lambda x.a},
%     {(X\appsep a),a}}}
%   {{{A,\lambda x.a},
%     {(A\appsep a),a}}}
%   {{{X,A,0}}}
%   {{}}

\noindent
We want to support this scenario but we have to improve a little our compiler
and \progress routines.
In particular $\hoUnifPb_1$ generates $\sigma = \{ A
\mapsto \lambda x. a\}$ making $\sigma\hoUnifPb_2$ equal to
$(\lambda x. a) \appsep a \Ue a$ that \Ue{} cannot solve since it
lacks rules \ref{clause:beta1} and \ref{clause:beta2}.

To address this problem the compiler must recognize and replace
\notllambda terms with fresh variables and generate a new kind of links
that we call \linkbeta.

In addition to \fullref{inv:linklhs}, the term on the \rhs of a \linkbeta
has the following property:

\newcommand{\rhsBetaHead}{\ensuremath{X_{s_1\dots s_n}}}
\newcommand{\rhsBeta}{\ensuremath{\rhsBetaHead\appsep t_1\dots t_m}\xspace}
% \noteE[inline]{Io direi $X_{s_1\ldots s_n}\appsep l_1\ldots l_m$
% ... where S = [s1 .. sn] e L = [l1 .. lm]}

\begin{invariant}[\linkbeta \rhs]
  The \rhs of any \linkbeta has the shape $X\appsep t_1\dots t_m$
  where $X$ is a unification variable in \llambda (with scope $s_1\dots s_n$)
  and $t_1\dots t_m$ is a list of terms such that $m>0$ and
  $t_1$ is either a variable occurring in $s_1\dots s_n$ or
  a term other than a variable.
  \label{inv:beta-rhs}
\end{invariant}

\noindent Note that the shape of such as \rhs is
  <<\elpiIn{app [uva X S | L]}>>, where \elpiIn{S}~$= s_1\dots s_n$
  and \elpiIn{L}~$=t_1\dots t_m$.

% \linkbeta are put in \linkStore and activated when \rhs falls in \llambda. 
% \noteE[inline]{un po' ridondante con 8.2, non so se serva}

% This property can be relaxed, to accept approximations: a dedicated
% section is given in a future section

% From \cref{lemma:keep-fidelity}, we can deduce the following corollary

% \begin{corollary}
%   Given a \linkbeta \linkbetaM{\Gamma}{X}{T}, if it exists a second link
%   $l\in\linkStore$, then unification fails.   
% \end{corollary}

\subsection{Compilation and decompilation}

For the detection of \notllambda it
is sufficient to identify applications with a flexible head and arguments that
are not in distinct names. The following rule for \notllambda compilation
is inserted  just before rule~\ref{rule:compapp}.

\input{code/comp_beta}

\noindent
The list \elpiIn{Ag} is split into the two parts: 
\elpiIn{Pf} is in \llambda, and can be empty; 
\elpiIn{Extra} cannot be empty and is such that
\elpiIn{append Pf Extra Ag}.
The \rhs of the \linkbeta is
the application of a fresh variable \elpiIn{C} having
in scope all names in \elpiIn{Pf1} (the compilation of \elpiIn{Pf}).
The variable \elpiIn{B}, returned has the compiled term, is a fresh variable
having in scope all the free variables occurring in \elpiIn{Pf1} and
\elpiIn{Extra1} (the compilation of \elpiIn{Extra}).
This construction enforces \cref{inv:beta-rhs}.

% In the following, we pose $X_{s_1\dots s_n}$ to represent a variable $X$ with
% the list of distinct names $s_1\dots s_n$ as scope. Moreover, $X_{s_1\dots
% s_n}\appsep t_1\dots t_m$ is the application of that variable to the list of
% terms $t_1\dots t_m$. Note that \rhsBeta is equivalent to \elpiIn{app[uva N S |
% T]}.

% \noteD{Si puo togliere se serve spazio}
% \begin{corollary}
%   Let \rhsBeta be the \rhs of a \linkbeta, then $m > 0$.
% \end{corollary}

% \begin{comment}
% \begin{proof}[Proof sketch]
%   Assume we have a \linkbeta, by contradiction, if $m = 0$, then the original
%   \Fo{} term has the shape \elpiIn{fapp[fuva M | Ag]} where \elpiIn{Ag} is a
%   list of distinct names (i.e. the list \elpiIn{Extra} is empty). This case is
%   however captured by rule~\ref{rule:complam} (from \cref{sec:compilation}) and
%   no \linkbeta is produced which contradicts our initial assumption.
% \end{proof}
% \end{comment}

% \noteD{Si puo togliere se serve spazio}
% \begin{corollary}
%   Let \rhsBeta be the \rhs of a \linkbeta, then $t_1$ either appears in $s_1\dots s_n$ or it is 
%   not a name.
% \end{corollary}

% \begin{comment}
% \begin{proof}[Proof sketch]
%   By construction, the lists $s_1\dots s_n$ and $t_1\dots t_m$ are built by splitting
%   the list \elpiIn{Ag} from the original term \elpiIn{fapp [fuva A|Ag]}.
%   $s_1\dots s_n$ is the longest prefix of the compiled terms in \elpiIn{Ag} which is
%   in \llambda. Therefore, by definition of \llambda, $t_1$ must appear  
%   in $s_1\dots s_n$, otherwise $s_1\dots s_n$ is not the longest prefix in
%   \llambda, or it is a term with a constructor of \elpiIn{tm} as functor.  
% \end{proof}
% \end{comment}

\paragraph{Decompilation}
All \linkbeta should be solved before decompilation.
If any \linkbeta remains in \linkStore, decompilation fails.

\subsection{Progress}

\newcommand{\progBetaLL}{\emph{\llambda-progress-refine}\xspace}
\newcommand{\progBetaRH}{\emph{\llambda-progress-\rhs}\xspace}
% \newcommand{\progBetaDedup}{\emph{progress-\llambda-dedup}\xspace}
\newcommand{\progBetaFail}{\emph{\llambda-progress-fail}\xspace}

Given a \linkbeta $l$ of the form \linkbetaM{\Gamma}{T}{\rhsBeta}, we provide
the following activation rules:

\begin{definition}[\progBetaLL]
  Let $\sigma$ be a substitution such that $\sigma t_1$ is a name $s$
  not occurring$s_1\dots s_n$. If $m = 1$, then $l$ is removed and \lhs is
  unified with $X_{s_1\ldots s_n~s}$.
  \todo{I changed this part, it is OK?}
  If $m > 1$, then $l$ is replaced by the link
  $\linkbetaM{\Gamma}{T}{Y_{s_1\dots s_n~s}\appsep t_2\dots t_m}$
  (where $Y$ is a fresh variable of arity $n+1$) and link
  \linketaM{\Gamma}{X_{s_1\dots s_n}}{\lambda x.Y_{s_1\dots s_n~x}}
  is added to \linkStore.

  % A link \linkbetaM{\Gamma}{T}{\rhsBeta} is removed from \linkStore, if given
  % the substitution $\sigma$, $\sigma t_1$ is a name, say $t$, such that, $t
  % \notin s_1\dots s_n$. In this case, let $Y$ a fresh variable, then
  % \linketaM{\Gamma}{X_{s_1\dots s_n}}{\lambda x.Y_{s_1\dots s_n,x}} is added to
  % \linkStore, and 1) if $m = 1$ then $Y_{s_1\dots s_n, t}$ is unified with $T$,
  % 2) otherwise, the refined \linkbeta, \linkbetaM{\Gamma}{T}{M_{s_1\dots
  % s_n,t}\appsep t_2\dots t_m}, is added to \linkStore.
  \label{def:progBetaLL}
  % % \noteE[inline]{non è chiaro. la sostituzione immagino sia la corrente, non se ne esiste una. Poi non è chiaro se otherwise si riferische a in this case. Usa un enumerate}
  % %\noteE[inline]{Is it clearer?}
  % \noteE[inline]{l1 or t1? Forse è più chiaro dire che il ilnk beta viene raffinato con link beta cone meno argomenti, Se poi il numero di argomenti extra è 0 il tutto è in llambda e quindi viene buttato via. In questo modo la terminazione è ancora più chiara perchè si vede già che prima decresce la lista di argomenti e poi il numero di beta}
  % \noteD{L'ho riformulato}
\end{definition}

\begin{definition}[\progBetaRH]
  Link $l$ is removed from
  \linkStore if \rhsBetaHead is instantiated to a term $t$ and
  $t \appsep t_1 \ldots t_m$ $\beta$-reduces to a $t' \in \llambda$.
  \label{def:progBetaRH}
\end{definition}

% Example of this `X a = Y a'
% \begin{definition}[\progBetaDedup]
%   Given a \linkbeta $l_1$ and second link $l_2 \in\linkStore$, such that they
%   share the same \lhs. The two \rhs are unified and $l_2$ is
%   removed from \linkStore.
%   \label{def:progBetaDedup}
% \end{definition}

\begin{definition}[\progBetaFail]
  \progress fails whenever:
  there exists another link $l' \in \linkStore$ with same \lhs as $l$;
  or the \lhs  of $l$ become rigid.
  \label{def:progBetaFail}
\end{definition}

\noindent In order to accommodate for the implementation of heuristics we
shall relax these failure conditions in~\cref{sec:heuristics}.

\begin{lemma}
  \elpiIn{progress} terminates
\end{lemma}

\begin{proof}[Proof sketch]
  Let $l$ a \linkbeta in the store \linkStore. If $l$ is activated by
  \progBetaRH{}, then it disappears from \linkStore and \elpiIn{progress}
  terminates. Otherwise, the \rhs of $l$ is made by a variable applied to $m$
  arguments. At each activation of \progBetaLL, $l$ is replaced by a new
  \linkbeta $l^1$ having $m-1$ arguments. At the $m^{th}$ iteration, the
  \linkbeta $l^m$ has no more arguments and is removed from \linkStore.
  Note that at the $m^{th}$ iteration, $m$ new \linketa have been added to
  \linkStore, however, by \cref{lemma:prog-eta-terminates}, the algorithm
  terminates. Finally \progBetaFail also guarantees termination since 
  it makes \elpiIn{progress} immediately fails.

  % \Cref{def:progBetaRH} makes \elpiIn{progress} terminates, since it makes a
  % \linkbeta disappear from \linkStore. On the other hand, \cref{def:progBetaLL},
  % creates each time a new \linketa and replace the old \linkbeta with a new one.
  % This progression can however triggered at most $m$ times, since each new
  % \linkbeta as a smaller list of applied variables. At the $m^{th}$ progression,
  % the \linkbeta is removed from \linkStore which is now filled by $m$ new
  % \linketa. By \cref{lemma:prog-eta-terminates}, we know that \elpiIn{progress}
  % terminates if \linkStore is made by only \linketa and therefore
  % \elpiIn{progress} still terminates.
  %\noteE[inline]{is it ok?}
  % \noteE[inline]{funziona. per essere più precisi io parlerei di ordine lessicografico (tipico ordine ben fondato usato per dimostrare terminazione). Nl nostro caso è la tripla (argomenti extra dei beta, numero di beta, numero di eta).}
\end{proof}

% \begin{corollary}
%   Given a \linkbeta, the variables occurring in its \rhs are in \llambda.
%   % \noteE[inline]{vuoi dire: le variabili che occorrono nel rhs sono in...?}
%   \noteD[inline]{is it clearer?}
% \end{corollary}

% \begin{proof}[Proof sketch]
%   By construction, the \rhs of \linkbeta has the shape \rhsBeta, $s_1\dots s_n$
%   is in \llambda and all the terms $t_1\dots t_n$ are in \llambda, too. If a
%   \linkbeta is triggered by \progBetaRH, then, by \cref{def:progBetaRH}, that
%   link is removed by \linkStore, and the property is satisfied. If the \linketa
%   is activated by \progBetaLL, then, by \cref{def:progBetaLL}, the new \linkbeta
%   as a variable as a scope which is still in \llambda. 
% \end{proof}

\begin{theorem}[Fidelity with \linkbeta]
  The introduction of \linkbeta guarantees \fullref{prop:fidelity-recovery}
  if \Uo{} fails outside \llambda.
\end{theorem}

\begin{proof}[Proof sketch]
  Let $\foUnifPb_i$ be the first problem in \foUnifPb not in \llambda and
  le $\sigma$ be the substitution obtained solving $\hoUnifPb_1\ldots\hoUnifPb_{i-1}$.
  If $\sigma\hoUnifPb_i$ is in \llambda, then each variable that the
  compiler used to replace a problemat term is assigned and hence
  by \cref{def:progBetaLL,def:progBetaRH} the associated \linkbeta
  been solved and removed. Note that in this case all
  calls to \Ue{} are between terms in \llambda and by \cref{lemma:fidelity-maybeeta}
  fidelity is guaranteed. If $\sigma\hoUnifPb$ is still in \notllambda,
  then by \cref{def:progBetaFail} unification fails as the
  corresponding unification in \Fo{} would.
\end{proof}

\paragraph{Example of \progBetaLL}
Consider the \linkbeta below:

% make test ONLY=7004 TEX=tex
\printAlll
  {{{X,\lambda x.x},
    {\lambda x.(Y\appsep (X\appsep x)),f}}}
  {{{A,\lambda x.x},
    {B,f}}}
  {{{Y,D,0},
    {X,A,0}}}
  {{{\eta,,A,\lambda x.E_{x}},
    {\eta,,B,\lambda x.C_{x}},
    {\llambda,x,C_{x},(D\appsep E_{x})}}}

\noindent
Initially the \linkbeta \rhs is a variable $D$ applied to the $E_x$.
The first unification problem results in $\sigma =
\{A \mapsto \lambda x.x\}$. In turn this instantiation
triggers $\linkStore_1$ by \progressetaleft 
and $E_x$ is assigned to $x$.
Under this substitution the \linkbeta becomes
\linkbetaM{x}{C_x}{(D \appsep x)}, and by \progBetaLL
it is replaced with the link:
\linketaM{}{E}{\lambda x.D_x}, while $C_x$ is unified with $D_x$. The second unification
problem assigns $f$ to $B$, that in turn activates
the second \linketa ($f$ is assigned to $C$), and then all the remaining links
are solved. The final \Ho substitution is $\sigma = \{~A _{} \mapsto \lambda x.x ~;~
B _{} \mapsto f ~;~ C _{x} \mapsto (f\appsep x) ~;~
D _{} \mapsto f ~;~ E _{x} \mapsto x ~;~
F _{x} \mapsto C_{x} ~\}$ and is decompiled into $\rho = \{~X \mapsto \lambda x.x ~;~
Y \mapsto f~\}$.

\paragraph{Example of \progBetaRH}
We give a second look at the opening example in \cref{sec:beta}.
\todo{this is a little more complex that the opening problem}
The problem is compiled into:
%
% make test ONLY=7003 TEX=tex
\printAlll
  {{{X,\lambda x.Y},
    {(X\appsep a),a}}}
  {{{A,\lambda x.B},
    {C,a}}}
  {{{Y,B,0},
    {X,A,0}}}
  {{{\llambda,,C,(A\appsep a)}}}

% \todo{Rinominare i print dei problemi}

The first unification problems is solved by the substitution
$\sigma = \{~A \mapsto \lambda x.B~\}$. The \linkbeta becomes
\linkbetaM{}{C}{((\lambda x.B) \appsep a)} whose \rhs can be $\beta$-reduced to
$B$. $B$ is in \llambda and is unified with $C$. The resolution of the second
unification problem
gives the final substitution $\sigma = \{~A \mapsto \lambda x.B ~;~ B \mapsto C ~;~ C
\mapsto a ~\}$ that is decompiled into $\rho = \{~ X \mapsto \lambda x.a ~;~ Y \mapsto
a ~\}$.

\subsection{\texorpdfstring{Relaxing \fullref{def:progBetaFail}}{Relaxing progress-llam-fail}}
\label{sec:heuristics}
\newcommand{\progBetaNoLLWait}{\emph{progress-beta-\notllambda}}

% from https://www.cs.mcgill.ca/~bpientka/papers/unif_miller60.pdf
Working with terms in \llambda is sometime too restrictive~\cite{Abel2018ExtensionsTM}
and we could find in literature a few strategies to go beyond \llambda without 
implementing Huet's algorithm~\cite{Huet75}.
Some implementations of $\lambda$Prolog~\cite{lamProlog} such as
Teyjus~\cite{Nadathur2001} delay the resolution of \notllambda unification
problems until the substitution makes them reenter \llambda. Other systems, for
example of the unification algorithm of Coq used in its type-class
solver~\cite{sozeau08}, apply heuristics like preferring projection over mimic,
and commit to that solution.

In this section we show how we can implement these strategies
by simply adding (or removing) rules to the
\elpiIn{progress} predicate.
In the example below $\foUnifPb_1$ is in \notllambda.
%
% make test ONLY=7005 TEX=tex
\printAlll
  {{{(X\appsep a),a},
    {X,\lambda x.Y}}}
  {{{A,a},
    {B,\lambda x.C}}}
  {{{Y,C,0},
    {X,B,0}}}
  {{{\llambda,,A,(B\appsep a)}}}

\noindent
If we want the object language unification to delay the first unification problem
(waiting for $X$ to be be instantiated),
we can relax \cref{def:progBetaFail}. 
Instead of failing when the \lhs of $\linkStore_1$ becomes rigid (equal to $a$),
we keep it in \linkStore until the head of its \rhs also become rigid.
In this case, since both the \lhs and \rhs have rigid heads, they can be unified.
While this relaxed rule does not break \fullref{prop:fidelity} per se,
the \elpiIn{occur-check-links} procedure becomes incomplete since
\fullref{inv:linklhs} is broken. Also note that delaying unification outside \llambda can leave \linkbeta for the
decompilation phase. Therefore \elpiIn{commit-links} should
be modified accordingly.

If instead we want \Uo{} to follow the second strategy and pick an
arbitrary solution we can modify \progress by applying the desired heuristic.
For instance, in $X \appsep a \appsep b = Y \appsep b$, the last
argument of the two terms is the same and unification can succeed by assigning
$X a$ to $Y$. This heuristic can be implemented by adding the following rule:

\input{code/progress_heuristic}

\noindent
Note that \elpiIn{UV} is in \llambda as well as \elpiIn{T}.

\section{Actual implementation in Elpi}\label{sec:implementation}

In this paper we show a minimized example. The full code is there.
But we also have to code things in Coq-Elpi.

The main difference between the presentation in the previous sections and
the actual implementation for Coq is that the main loop \hrun is replaced by
the one of Prolog that chains calls to the unification procedure. In order
implement the store of links we resort to Elpi's CLP engine and
use constraints (suspended goals) to represent links, and constraint
handling rules to implement progress operations involving more than one link.

\todo{finish}
about the progress of 1 link:

\begin{elpicode}
link-eta L R :- suspend-condition L R Holes, !,
  declare_constraint (link-eta L R) Holes.
link-eta L R :-
  progress. % e.g. L = R.
\end{elpicode}

about the progress of 2 links:

\begin{elpicode}
constraint link-eta {
  rule (N1 ~$\triangleright$~ G1 ?- link-eta (uvar X LX1) T1) % match
    /  (N2 ~$\triangleright$~ G2 ?- link-eta (uvar X LX2) T2) % remove
    |  (relocate LX1 LX2 T2 T2')             % condition
   <=> (N1 ~$\triangleright$~ G1 ?- T1 = T2').                % new goal
}
\end{elpicode}

Remark how the invariant about uvar arity makes this easy, since LX1 and LX2
have the same length. Also note that N1 only contains the names of the first
link (while relocate runs in the disjoint union) and Elpi ensures that
T2' can live in N1.

\noindent
\section{Related work and conclusion}

Different strategies can be used to unify terms of the object language.
% in the meta langauge. 
The first approach that comes to mind consist in implementing \Uo as a
regular routine in the ML as follows:

\begin{elpicode}
  decision X :- X ~\Uo~(all A x\ app [P, x]), finite A,
    pi x\ decision (app [P, x]).
\end{elpicode}

\noindent
Opting for this method would result in a suboptimal utilization of the logic
programming engine provided by the meta language, as it degrades indexing by
eliminating all data from rule heads. Additionally,
implementing an unification procedure in the meta
language is likely to be significantly slower compared to the built-in one.

% Paper \cite{10.1145/2966268.2966272} introduces semi-shallow.
Another possibility is to avoid having application and abstraction nodes
in the syntax tree, and use the ones of meta language, as in the following:

\begin{elpicode}
decision (all A x\ P x) :- finite A, pi x\ decision (P x).
\end{elpicode}

\noindent
However, this encoding has two big limitations. Firstly, in CIC, it is
not always feasible to adopt it due to the meta language's type system being
too limited to accommodate for that one the object language. E.g. CIC
can typecheck variadic functions, see for example~\ref{CPDT chlipala}. \\
Secondly, the CIC encoding provided by Elpi is primarily utilized for meta
programming, in order to extend the Coq system. Consequently, it must be able to
manipulate terms that are not known in advance
without relying on introspection primitives such as Prolog's
\texttt{functor} and \texttt{arg}. In this context, constants need to live in
an open world, akin to the \elpiIn{string} data type used in the preceding
examples.

In the literature we could find a related encoding of the Calculus of
Constructions (CC)~\cite{felty93lics}. The goal of that work was to exhibit
a logic program performing proof checking for CC and hence relate the
proof system of intuitionistic higher-order logic (that animates $\lambda$Prolog
programs) with the one of CC. The encoding is hence tailored
toward a different goal, for example it utilizes three relations to represent the
equational theory of CC. Section 6 contains a discussion about the use of the
unification procedure of the meta language in presence of non ground goals, but
the authors do not aim at exploiting it to the degree we want.\todo{sucks}

\todo{cite isabelle's TC, that are baked in}

\emph{The approach presented in this paper}
provides a third option that addresses all the concerns
mentioned earlier. It capitalizes on the benefit of not requiring to fully
implement
the unification algorithm of the object language. Instead, it employs the
unification capabilities of the meta language, facilitated by the various links
%we establish
to manage ``problematic'' subterms. As a result of this choice our encoding takes
advantage of indexing data structures and mode analysis for clause filtering.
It it worth mentioning that we only replace terms with variables only
when it is strictly needed, leaving the rest of the term structure intact
and hence indexable by the meta language logic programming engine.
Moreover, the unification process we obtain is ready to take
advantage of potential improvements to the programming engine,
such as tabled search, and apply forms of static analysis for
the meta language, such as determinacy, to the object language. 
Finally, our approach is flexible enough to accommodate different strategies
and heuristics to handle terms outside the pattern fragment, and
it is not tightly coupled with CIC.

% ---

% Benefits: less work, reuse efficient ho unif (3x faster), indexing,

% Future: tabling and static analysis (reuse for ML again).

% Very little is Coq specific. Applies to all OL that are not a subsystem of 
% HOL, or for ML that are used for meta programming.

\printbibliography

\clearpage

\input{appendix.tex}

\end{document}