### 2024-07-24 paper/paper_for_taps/source/code/meta_language.tex

\begin{elpicode}

  kind inctx type -> type.                              ~($\customlabel{data:inctx}{\cdot\vdash\cdot}$)~
  type abs (tm -> inctx A) -> inctx A.
  type val A -> inctx A.
  typeabbrev assignment (inctx tm).
  typeabbrev subst (mem assignment).

  kind tm  type.
  type app list tm -> tm.
  type lam (tm -> tm) -> tm.
  type con string -> tm.
  type uva addr -> list tm -> tm.

  ~\PYG{k+kd}{type} \PYG{n+nf}{(\Ue)} \PYG{k+kt}{tm -> tm -> subst -> subst -> o}~.
  (con C ~$\Ue$~ con C) S S.
  (app L1 ~$\Ue$~ app L2) S S1 :- fold2 (~$\Ue$~) L1 L2 S S1.
  (lam F1 ~$\Ue$~ lam F2) S S1 :- 
    pi x\ (pi S\ (x ~$\Ue$~ x) S S) => (F1 x ~$\Ue$~ F2 x) S S1.
  (uva N Args ~$\Ue$~ T) S S1 :- 
    set? N S F,!, move F Args T1, (T1 ~$\Ue$~ T) S S1.
  (T ~$\Ue$~ uva N Args) S S1 :- 
    set? N S F,!, move F Args T1, (T ~$\Ue$~ T1) S S1.
  (uva M A1 ~$\Ue$~ uva N A2) S1 S2 :- !, 
    pattern-fragment A1, pattern-fragment A2, 
    prune! M A1 N A2 S1 S2.
  (uva N Args ~$\Ue$~ T) S S1 :- not_occ N S T, pattern-fragment Args, 
    bind T Args T1, assign N S T1 S1.
  (T ~$\Ue$~ uva N Args) S S1 :- not_occ N S T, pattern-fragment Args, 
    bind T Args T1, assign N S T1 S1.

  type prune! addr -> list tm -> addr -> 
              list tm -> subst -> subst -> o.
  /* no pruning needed */               
  prune! N A  N A  S S :- !.
  prune! M A  N A  S1 S2 :- !, bind (uva M A) A Ass, 
    assign N S1 Ass S2.
  /* prune different arguments */
  prune! N A1 N A2 S1 S3 :- !,
    new S1 W S2, prune-same-variable W A1 A2 [] Ass,
    assign N S2 Ass S3.
  /* prune to the intersection of scopes */
  prune! N A1 M A2 S1 S4 :- !,
    new S1 W S2, prune-diff-variables W A1 A2 Ass1 Ass2, 
    assign N S2 Ass1 S3,
    assign M S3 Ass2 S4.

  type prune-same-variable addr -> list tm -> list tm -> 
                            list tm -> assignment -> o.
  prune-same-variable N [] [] ACC (val (uva N Args)) :- 
    rev ACC Args.
  prune-same-variable N [X|XS] [X|YS] ACC (abs F) :- 
    pi x\ prune-same-variable N XS YS [x|ACC] (F x).
  prune-same-variable N [_|XS] [_|YS] ACC (abs F) :- 
    pi x\ prune-same-variable N XS YS ACC (F x).

  type permute list nat -> list tm -> list tm -> o.
  permute [] _ [].
  permute [P|PS] Args [T|TS] :-
    nth P Args T,
    permute PS Args TS.

  type build-perm-assign addr -> list tm -> list bool -> 
                      list nat -> assignment -> o.
  build-perm-assign N ArgsR [] Perm (val (uva N PermutedArgs)) :- 
    rev ArgsR Args, permute Perm Args PermutedArgs.
  build-perm-assign N Acc [tt|L] Perm (abs T) :- 
    pi x\ build-perm-assign N [x|Acc] L Perm (T x).
  build-perm-assign N Acc [ff|L] Perm (abs T) :- 
    pi x\ build-perm-assign N Acc L Perm (T x).

  type keep list A -> A -> bool -> o.
  keep L A tt :- mem L A, !.
  keep _ _ ff.

  type prune-diff-variables addr -> list tm -> list tm -> 
                            assignment -> assignment -> o.
  prune-diff-variables N Args1 Args2 Ass1 Ass2 :-
    map (keep Args2) Args1 Bits1, 
    map (keep Args1) Args2 Bits2,
    filter Args1 (mem Args2) ToKeep1,
    filter Args2 (mem Args1) ToKeep2,
    map (index ToKeep1) ToKeep1 IdPerm,
    map (index ToKeep1) ToKeep2 Perm21,
    build-perm-assign N [] Bits1 IdPerm Ass1,
    build-perm-assign N [] Bits2 Perm21 Ass2.

  type beta tm -> list tm -> tm -> o.
  beta A [] A :- !.
  beta (lam Bo) [H | L] R :- beta (Bo H) L R1, beta-aux R1 R.
  beta (app A) L (app X) :- append A L X.
  beta (con H) L (app [con H | L]).
  beta X L (app[X|L]) :- name X.

  type beta-aux tm -> tm -> o.
  beta-aux (app [HD|TL]) R :- !, beta HD TL R.
  beta-aux A A.

  /* occur check for N before crossing a functor */
  type not_occ  addr -> subst -> tm -> o.
  not_occ N S (uva M Args) :- set? M S F, 
    move F Args T, not_occ N S T.
  not_occ N S (uva M Args) :- unset? M S, not (M = N),
    forall1 (not_occ_aux N S) Args.
  not_occ _ _ (con _).
  not_occ N S (app L) :- not_occ_aux N S (app L).
  /* Note: lam is a functor for the meta language! */
  not_occ N S (lam L) :- pi x\ not_occ_aux N S (L x).
  not_occ _ _ X :- name X.
  /* finding N is ok */
  not_occ N _ (uva N _).

  /* occur check for X after crossing a functor */
  type not_occ_aux addr -> subst -> tm -> o.
  not_occ_aux N S (uva M _) :- unset? M S, not (N = M).
  not_occ_aux N S (uva M Args) :- set? M S F, 
    move F Args T, not_occ_aux N S T.
  not_occ_aux N S (app L) :- forall1 (not_occ_aux N S) L.
  not_occ_aux N S (lam F) :- pi x\ not_occ_aux N S (F x).
  not_occ_aux _ _ (con _).
  not_occ_aux _ _ X :- name X.
  /* finding N is ko, hence no rule */

  /* copy T T' vails if T contains a free variable, i.e. it
     performs scope checking for bind */
  type copy tm -> tm -> o.
  copy (con C)   (con C).
  copy (app L)   (app L') :- map copy L L'.
  copy (lam T)   (lam T') :- pi x\ copy x x => copy (T x) (T' x).
  copy (uva A L) (uva A L') :- map copy L L'.

  type bind tm -> list tm -> assignment -> o.
  bind T [] (val T') :- copy T T'.
  bind T [X | TL] (abs T') :- pi x\ copy X x => bind T TL (T' x).

  type deref subst -> tm -> tm -> o.                   ~($\customlabel{pred:deref}{\sigma t}$)~
  deref _ (con C) (con C).
  deref S (app A) (app B) :- map (deref S) A B.
  deref S (lam F) (lam G) :-
    pi x\ deref S x x => deref S (F x) (G x).
  deref S (uva N L) R :- set? N S A,
    move A L T, deref S T R.
  deref S (uva N A) (uva N B) :- unset? N S,
    map (deref S) A B.

  type move assignment -> list tm -> tm -> o.
  move (abs Bo) [H|L] R :- move (Bo H) L R.
  move (val A)  [] A :- !.

  type deref-assmt subst -> assignment -> assignment -> o.
  deref-assmt S (abs T) (abs R) :- pi x\ deref-assmt S (T x) (R x).
  deref-assmt S (val T) (val R) :- deref S T R.
\end{elpicode}