### 2024-07-26 paper/paper_for_taps/source_src/code/common.tex
new version (seems accepted by taps)

\begin{elpicode}
kind nat type.
type z nat.
type s nat -> nat.

kind bool type.
type tt bool.
type ff bool.

kind option type -> type.
type some A -> option A.
type none option A.

type int->nat int -> nat -> o.
int->nat 0 z.
int->nat R (s N) :- R > 0, R' is R - 1, int->nat R' N.

type nth nat -> list A -> A -> o.
nth z (X :: _) Y  :- !, X = Y.
nth (s N) (_ :: L) X :- nth N L X.

type list-init nat -> nat -> (nat -> A -> o) -> list A -> o.
list-init Len Len _ [].
list-init Len Pos F [A | TL] :- F Pos A, list-init Len (s Pos) F TL.

type iota nat -> list nat -> o.
iota N L :- list-init N z (=) L.

type partition (A -> o) -> list A -> list A -> list A -> o.
partition _ [] [] [].
partition F [A|As] [A|Bs] Cs :- F A, !, partition F As Bs Cs.
partition F [A|As] Bs [A|Cs] :- partition F As Bs Cs.

type index list A -> A -> nat -> o.
index [A|_] A z :- !.
index [_|TL] A (s P) :- index TL A P.

type take-list list A -> A -> list A -> o.
take-list [L|XS] L XS.
take-list [L|XS] L' [L|YS] :- not(L = L'), take-list XS L' YS.

type not-in list A -> A -> o.
not-in [] _.
not-in [X | TL] T :- not (X = T), not-in TL T.

type exists (A -> o) -> list A -> o.
exists F [X|XS] :- F X ; exists F XS.

type forall2 (A -> B -> o) -> list A -> list B -> o.
forall2 _ [] [].
forall2 F [X|XS] [Y|YS] :- F X Y, forall2 F XS YS.

type forall3 (A -> B -> C -> o) -> list A -> list B -> list C -> o.
forall3 _ [] [] [].
forall3 F [X|XS] [Y|YS] [Z|ZS] :- F X Y Z, forall3 F XS YS ZS.

/* same as forall2, sometimes more intuitive name */
type map (A -> B -> o) -> list A -> list B -> o.
map _ [] [].
map F [X|XS] [Y|YS] :- F X Y, map F XS YS.

type fold2 (A -> A1 -> B -> B -> o) -> list A -> list A1 -> B -> B -> o.
fold2 _ [] [] A A.
fold2 F [X|XS] [Y|YS] A A1 :- F X Y A A0, fold2 F XS YS A0 A1.

type fold (A -> B -> B -> o) -> list A -> B -> B -> o.
fold _ [] A A.
fold F [X|XS] A A1 :- F X A A0, fold F XS A0 A1.

~\PYG{k+kd}{type} \PYG{n+nf}{(>n)} \PYG{k+kt}{nat -> nat -> o}~.
s _ >n z.
s N >n s M :- N >n M.

~\PYG{k+kd}{type} \PYG{n+nf}{(>=n)} \PYG{k+kt}{nat -> nat -> o}~.
_ >=n z.
s N >=n s M :- N >=n M.

type diff nat -> nat -> o.
diff N M :- N >n M ; M >n N.

type map-fold2 (A -> A1 -> B -> B -> C -> C -> o) -> list A -> B -> C -> list A1 -> B -> C -> o.
map-fold2 _F [] B C [] B C.
map-fold2 F [X | XS] B C [Y | YS] B2 C2 :-
  F X Y B B1 C C1,
  map-fold2 F XS B1 C1 YS B2 C2.

type mem (list A) -> A -> o.
mem [A|_] A :- !.
mem [_|L] A :- mem L A.

type append list A -> list A -> list A -> o.
append [] A A.
append [A | TL] L [A | L'] :- append TL L L'.

type fold4 (A -> A1 -> B -> B -> C -> C -> o) -> list A -> list A1 -> B -> B -> C -> C -> o.
fold4 _ [] [] A A B B.
fold4 F [X|XS] [Y|YS] A A1 B B1 :- F X Y A A0 B B0, fold4 F XS YS A0 A1 B0 B1.

type fold6 (A -> A1 -> B -> B -> C -> C -> D -> D -> o) -> list A -> list A1 -> B -> B -> C -> C -> D -> D -> o.
fold6 _ [] [] A A B B C C.
fold6 F [X|XS] [Y|YS] A A1 B B1 C C1 :- F X Y A A0 B B0 C C0, fold6 F XS YS A0 A1 B0 B1 C0 C1.

type flatten-aux list A -> list (list A) -> list A -> o.
flatten-aux [] L L1 :- flatten L L1.
flatten-aux [A|As] L [A|L1] :- flatten-aux As L L1.

type flatten list (list A) -> list A -> o.
flatten [] [].
flatten [H|L] R :- flatten-aux H L R.

type all list o -> o.
all [].
all [P|PS] :- P, all PS.

type forall1 (X -> o) -> list X -> o.
forall1 _ L :- var L, !.
forall1 P [X|XS] :- P X, forall1 P XS.
forall1 _ [].

type none (X -> o) -> list X -> o.
none _ [].
none P [X|Xs] :- not (P X), none P Xs.

type undup list A -> list A -> o.
undup [] [].
undup [X|Xs] [X|R] :- not-in Xs X, !, undup Xs R.
undup [_|Xs] R :- undup Xs R.

type len list A -> nat -> o.
len [] z.
len [_|L] (s X) :- len L X.

type rev list A -> list A -> o.
rev L RL  :- rev.aux L []  RL.
rev.aux [X|XS] ACC R :- rev.aux XS [X|ACC] R.
rev.aux [] L L.

type filter list A -> (A -> prop) -> list A -> o.
filter []    _ [].
filter [X|L] P R :- if (P X) (R = X :: L1) (R = L1), filter L P L1.

type filter2 list A -> list B -> (A -> B -> prop) -> list A -> o.
filter2 [] []   _ [].
filter2 [X|L] [Y|YS] P R :- if (P X Y) (R = X :: L1) (R = L1), filter2 L YS P L1.

type pattern-fragment-prefix-aux list A -> list A -> list A -> list A -> o.
pattern-fragment-prefix-aux [] _ [] [].
pattern-fragment-prefix-aux [X | L] ACC [X | PF] NPF :- name X, not-in ACC X, !, 
  pattern-fragment-prefix-aux L [X | ACC] PF NPF.
pattern-fragment-prefix-aux L _ [] L.

type pattern-fragment-prefix list A -> list A -> list A -> o.
pattern-fragment-prefix L PF NPF :- pattern-fragment-prefix-aux L [] PF NPF.

type pattern-fragment list A -> o.
pattern-fragment L :- distinct_names L.
\end{elpicode}