### 2024-07-26 paper/paper_for_taps/source_src/code/maybe_eta.tex

\begin{elpicode}
  type occurs-rigidly to -> to -> o.
  occurs-rigidly N N.
  occurs-rigidly _ (o-app [o-uva _|_]) :- !, fail.
  occurs-rigidly N (o-app L) :- exists (occurs-rigidly N) L.
  occurs-rigidly N (o-lam B) :- pi x\ occurs-rigidly N (B x).

  type reducible-to list to -> to -> to -> o.
  reducible-to _ N N :- !.
  reducible-to L N (o-app[o-uva _|Args]) :- !, 
    forall1 (x\ exists (reducible-to [] x) Args) [N|L]. 
  reducible-to L N (o-lam B) :- !, 
    pi x\ reducible-to [x | L] N (B x).
  reducible-to L N (o-app [N|Args]) :-
    last-n {len L} Args R,
    forall2 (reducible-to []) R {rev L}.

  type maybe-eta to -> list to -> o.                  ~\customlabel{rule:maybeeta}{(\maybeeta)}~
  maybe-eta (o-app[o-uva _|Args]) L :- !,
    forall1 (x\ exists (reducible-to [] x) Args) L, !. 
  maybe-eta (o-lam B) L :- !, pi x\ maybe-eta (B x) [x | L].
  maybe-eta (o-app [T|Args]) L :- (name T; T = o-con _),
    split-last-n {len L} Args First Last,
    none (x\ exists (y\ occurs-rigidly x y) First) L,
    forall2 (reducible-to []) {rev L} Last.
\end{elpicode}