### 2024-07-26 paper/paper_for_taps/source_src/code/test.tex

\begin{elpicode}

type hstep tm -> tm -> links -> links -> subst -> subst -> o.
hstep T1 T2 L1 L2 S1 S3 :-
  (T1 ~$\Ue$~ T2) S1 S2,
  progress L1 L2 S2 S3.

type compile to -> tm -> mmap -> mmap ->  links -> links -> 
  subst -> subst -> o.
compile F G M1 M2 L1 L2 S1 S2 :-
  beta-normal F F', comp F' G M1 M2 L1 L2 S1 S2.

~\PYG{k+kd}{type} \PYG{n+nf}{(\Uo)} \PYG{k+kt}{to -> to -> fsubst -> o}~.
(A ~$\Uo$~ B) F :-
  compile A A' [] M1 [] L1 [] S1,
  compile B B' M1 M2 L1 L2 S1 S2,
  hstep A' B' L2 L3 S2 S3,
  decompile M2 L3 S3 [] F.

\end{elpicode}